### 2015-08-05 afidraftRnw.tex

\documentclass[a4paper,man,natbib,12pt,apacite]{apa6}
\usepackage{changepage}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{varioref}
\usepackage[longnamesfirst]{natbib}
%\usepackage[natbibapa]{apacite}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{lineno}
\usepackage{listings}
\usepackage{longtable}
\usepackage{verbatim}
\usepackage[yyyymmdd,hhmmss]{datetime}

\newlength{\wideitemsep}
\setlength{\wideitemsep}{\itemsep}
\addtolength{\wideitemsep}{-10pt}
\let\olditem\item
\renewcommand{\item}{\setlength{\itemsep}{\wideitemsep}\olditem}

% New definition of square root:
% it renames \sqrt as \oldsqrt
\let\oldsqrt\sqrt
% it defines the new \sqrt in terms of the old one
\def\sqrt{\mathpalette\DHLhksqrt}
\def\DHLhksqrt#1#2{%
\setbox0=\hbox{$#1\oldsqrt{#2\,}$}\dimen0=\ht0
\advance\dimen0-0.2\ht0
\setbox2=\hbox{\vrule height\ht0 depth -\dimen0}%
{\box0\lower0.4pt\box2}}
\makeatletter
\newread\pin@file
\newcounter{pinlineno}
\newcommand\pin@accu{}
\newcommand\pin@ext{pintmp}
% inputs #3, selecting only lines #1 to #2 (inclusive)
\newcommand*\partialinput [3] {%
  \IfFileExists{#3}{%
    \openin\pin@file #3
    % skip lines 1 to #1 (exclusive)
    \setcounter{pinlineno}{1}
    \@whilenum\value{pinlineno}<#1 \do{%
      \read\pin@file to\pin@line
      \stepcounter{pinlineno}%
    }
    % prepare reading lines #1 to #2 inclusive
    \addtocounter{pinlineno}{-1}
    \let\pin@accu\empty
    \begingroup
    \endlinechar\newlinechar
    \@whilenum\value{pinlineno}<#2 \do{%
      % use safe catcodes provided by e-TeX's \readline
      \readline\pin@file to\pin@line
      \edef\pin@accu{\pin@accu\pin@line}%
      \stepcounter{pinlineno}%
    }
    \closein\pin@file
    \expandafter\endgroup
    \scantokens\expandafter{\pin@accu}%
  }{%
    \errmessage{File `#3' doesn't exist!}%
  }%
}
\makeatother
\npdecimalsign{.}
\newcommand{\cil}[1]{\todo[inline, color=green!40]{#1}}
\newcommand{\cm}[1]{\todo[color=green!40]{#1}}
\newcommand{\cmb}[1]{\todo[color=blue!40]{\lstinline|#1|}}
\title{Intelligence and Age at First Intercourse: Cause or Confound?}
\shorttitle{Intelligence and AFI}
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}
\addtolength{\marginparwidth}{6cm}
\addtolength{\paperwidth}{\marginparwidth-1cm}
%\addtolength{\marginparsep}{2mm}
\abstract{Your abstract here.\\\today\ at \currenttime}
\authornote{This material is based upon work that has been supported by the National Science Foundation Graduate Research Fellowship Program under Grant No. (DGE-1445197) and various means of institutional support from the following universities: University of British Columbia \& Vanderbilt University.}
\begin{document}
\maketitle
%\linenumbers%\section{Introduction}\section{ }\vspace{-.8cm}
\cmb{content/introduction.txt}\input{content/introduction.txt}

\section{Present Study}
\begin{itemize}
\item We wanted to know whether the Harden \& Mendle (2011) results would hold using a different design and data set.\smallskip
\item We examined the relationship between AFI and Intelligence using siblings and their children from a multigenerational nationally representative sample.\smallskip
\item Specifically, we addressed the following questions:\smallskip
\begin{itemize}\item Does intelligence (mother or child) predict child AFI, after controlling for gene and environmental confounds?\smallskip
\item Is this relationship consistent between-- and within--families?
\end{itemize}\end{itemize}

\section{Methods}\cmb{content/methods.txt}\input{content/methods.txt}