### 2015-08-05 afidraftRnw.tex

\documentclass[a4paper,man,natbib,12pt,apacite]{apa6}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{changepage}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{varioref}
\usepackage[longnamesfirst]{natbib}
%\usepackage[natbibapa]{apacite}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{lineno}
\usepackage{listings}
\usepackage{longtable}
\usepackage{verbatim}
\usepackage[yyyymmdd,hhmmss]{datetime}

\newlength{\wideitemsep}
\setlength{\wideitemsep}{\itemsep}
\addtolength{\wideitemsep}{-10pt}
\let\olditem\item
\renewcommand{\item}{\setlength{\itemsep}{\wideitemsep}\olditem}

% New definition of square root:
% it renames \sqrt as \oldsqrt
\let\oldsqrt\sqrt
% it defines the new \sqrt in terms of the old one
\def\sqrt{\mathpalette\DHLhksqrt}
\def\DHLhksqrt#1#2{%
\setbox0=\hbox{$#1\oldsqrt{#2\,}$}\dimen0=\ht0
\advance\dimen0-0.2\ht0
\setbox2=\hbox{\vrule height\ht0 depth -\dimen0}%
{\box0\lower0.4pt\box2}}
\makeatletter
\newread\pin@file
\newcounter{pinlineno}
\newcommand\pin@accu{}
\newcommand\pin@ext{pintmp}
% inputs #3, selecting only lines #1 to #2 (inclusive)
\newcommand*\partialinput [3] {%
  \IfFileExists{#3}{%
    \openin\pin@file #3
    % skip lines 1 to #1 (exclusive)
    \setcounter{pinlineno}{1}
    \@whilenum\value{pinlineno}<#1 \do{%
      \read\pin@file to\pin@line
      \stepcounter{pinlineno}%
    }
    % prepare reading lines #1 to #2 inclusive
    \addtocounter{pinlineno}{-1}
    \let\pin@accu\empty
    \begingroup
    \endlinechar\newlinechar
    \@whilenum\value{pinlineno}<#2 \do{%
      % use safe catcodes provided by e-TeX's \readline
      \readline\pin@file to\pin@line
      \edef\pin@accu{\pin@accu\pin@line}%
      \stepcounter{pinlineno}%
    }
    \closein\pin@file
    \expandafter\endgroup
    \scantokens\expandafter{\pin@accu}%
  }{%
    \errmessage{File `#3' doesn't exist!}%
  }%
}
\makeatother
\npdecimalsign{.}
\newcommand{\cil}[1]{\todo[inline, color=green!40]{#1}}
\newcommand{\cm}[1]{\todo[color=green!40]{#1}}
\newcommand{\cmb}[1]{\todo[color=blue!40]{\lstinline|#1|}}
\title{Intelligence and Age at First Intercourse: Cause or Confound?}
\shorttitle{Intelligence and AFI}
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}
\addtolength{\marginparwidth}{6cm}
\addtolength{\paperwidth}{\marginparwidth-1cm}
%\addtolength{\marginparsep}{2mm}
\abstract{Your abstract here.\\\today\ at \currenttime}
\authornote{This material is based upon work that has been supported by the National Science Foundation Graduate Research Fellowship Program under Grant No. (DGE-1445197) and various means of institutional support from the following universities: University of British Columbia \& Vanderbilt University.}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\maketitle
%\linenumbers%\section{Introduction}\section{ }\vspace{-.8cm}
\cmb{content/introduction.txt}\input{content/introduction.txt}

\section{Present Study}
\begin{itemize}
\item We wanted to know whether the Harden \& Mendle (2011) results would hold using a different design and data set.\smallskip
\item We examined the relationship between AFI and Intelligence using siblings and their children from a multigenerational nationally representative sample.\smallskip
\item Specifically, we addressed the following questions:\smallskip
\begin{itemize}\item Does intelligence (mother or child) predict child AFI, after controlling for gene and environmental confounds?\smallskip
\item Is this relationship consistent between-- and within--families?
\end{itemize}\end{itemize}

\section{Methods}\cmb{content/methods.txt}\input{content/methods.txt}

\cmb{content/Gen1measure.txt}\input{content/Gen1measure.txt}

\cmb{content/Gen2measure.txt}\input{content/Gen2measure.txt}

\section{Results}\cmb{content/results.txt}\input{content/results.txt}
%\subsection{Between vs. Within Illustration}
%\cmb{content/results_meancomp.txt}\input{content/results_meancomp.txt}

\subsection{Between Family Anlayses}\cmb{content/results_between.txt}\input{content/results_between.txt}

\subsection{Within Family Analyses}
\cmb{content/results_between.txt}\input{content/results_within.txt}

\section{Discussion}

\newpage\section{Tables}\label{appen_tables}

\subsection{Descriptive Statistics}
\begin{longtable}{@{\extracolsep{5pt}}lcccccc}
\caption{Gen1 Summary Statistics by Sibling Status.}\label{table_summary_stats_sibinsampleg1}
\partialinput{2}{10}{table_summary_stats_sibinsampleg1.tex}
\end{longtable}\linebreak

\begin{longtable}{@{\extracolsep{5pt}}lcccccc}
\caption{Gen2 Summary Statistics by Sibling Status.}\label{table_summary_stats_sibinsampleg2}
\partialinput{2}{10}{table_summary_stats_sibinsampleg2.tex}
\end{longtable}\pagebreak
\npnoround
\begin{longtable}{@{\extracolsep{5pt}}lllcc}
\caption{Gen2 AFI by Gender, Race, and GenderxRace.}\label{table_afi_race_gender}
\partialinput{2}{18}{table_summary_stats_AFIRACEGENDER.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 9.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{table_g2_9measurement.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P.Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{table_g2loading_9.tex}
\end{longtable}\pagebreak

\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{24}{table_Mean_Mom_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{24}{table_Mean_Child_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{26}{table_Mean_Joint_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{28}{table_Dif_Mom_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{28}{table_Dif_Child_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{32}{table_Dif_Joint_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}

\begin{comment}\section{Figures}\label{figures}

\begin{figure}
  \centering
    \includegraphics[height=.85\textwidth,angle=270]{G1wrightMap}
     \caption{Wright Map of Gen1 Cognitive Ability}\label{G1wrightMap}
\end{figure}\end{comment}
\appendix\label{appen}
\section{Age 10.5 Replication}\label{appen10}
\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 10.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{table_g2_10measurement.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{table_g2loading_10.tex}
\end{longtable}\pagebreak
\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{24}{table_Mean_Mom_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{24}{table_Mean_Child_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{26}{table_Mean_Joint_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{28}{table_Dif_Mom_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{28}{table_Dif_Child_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{32}{table_Dif_Joint_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}

\section{Age 11.5 Replication}\label{appen11}

\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 11.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{table_g2_11measurement.tex}
\end{longtable}\pagebreak
\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P.Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{table_g2loading_11.tex}
\end{longtable}\pagebreak
\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{24}{table_Mean_Mom_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{24}{table_Mean_Child_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{26}{table_Mean_Joint_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{28}{table_Dif_Mom_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{28}{table_Dif_Child_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{32}{table_Dif_Joint_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}
%\section{Notes}
%\input{content/NOTES.txt}

\bibliography{AFI}
\end{document}