### 2015-08-28 afidraftRnw.tex
causes of afi

\documentclass[a4paper,man,apacite,natbib,12pt]{apa6}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
%%%%%%%%%%%% Packages %%%%%%%%%%%%
\usepackage{changepage}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[graphicx]{realboxes}
\usepackage{hyperref}

\usepackage[longnamesfirst]{natbib}
\usepackage[natbibapa,nodoi]{apacite}
\usepackage{enumitem}
\usepackage{lineno}
\usepackage{listings}
\usepackage{longtable}
\usepackage{verbatim}
\usepackage[yyyymmdd,hhmmss]{datetime}
\usepackage{pdflscape}
\usepackage{varioref} %references page location

%%%%%%%%%%%% New Commands %%%%%%%%%%%%

%%%Remove DOI %UNCOMMENT ALL to toggle between DOI and URL
\usepackage{etoolbox}
\usepackage{environ}

\newtoggle{bibdoi}
%\newtoggle{biburl}
\makeatletter

%\undef{\APACrefURL}
%\undef{\endAPACrefURL}
\undef{\APACrefDOI}
\undef{\endAPACrefDOI}

%\long\def\collect@url#1{\global\def\bib@url{#1}}
\long\def\collect@doi#1{\global\def\bib@doi{#1}}
%\newenvironment{APACrefURL}{\global\toggletrue{biburl}\Collect@Body\collect@url}{\unskip\unskip}
\newenvironment{APACrefDOI}{\global\toggletrue{bibdoi}\Collect@Body\collect@doi}{}

%\AtBeginEnvironment{thebibliography}{
% \pretocmd{\PrintBackRefs}{%
%  \iftoggle{bibdoi}
%    {\iftoggle{biburl}{\unskip\unskip doi:\bib@doi}{}}
%    {\iftoggle{biburl}{Retrieved from\bib@url}{}}
%  \togglefalse{bibdoi}\togglefalse{biburl}%
%  }{}{}
%}

%%% Partial Input
\usepackage[ascii]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\makeatletter
\newread\pin@file
\newcounter{pinlineno}
\newcommand\pin@accu{}
\newcommand\pin@ext{pintmp}
% inputs #3, selecting only lines #1 to #2 (inclusive)
\newcommand*\partialinput [3] {%
  \IfFileExists{#3}{%
    \openin\pin@file #3
    % skip lines 1 to #1 (exclusive)
    \setcounter{pinlineno}{1}
    \@whilenum\value{pinlineno}<#1 \do{%
      \read\pin@file to\pin@line
      \stepcounter{pinlineno}%
    }
    % prepare reading lines #1 to #2 inclusive
    \addtocounter{pinlineno}{-1}
    \let\pin@accu\empty
    \begingroup
    \endlinechar\newlinechar
    \@whilenum\value{pinlineno}<#2 \do{%
      % use safe catcodes provided by e-TeX's \readline
      \readline\pin@file to\pin@line
      \edef\pin@accu{\pin@accu\pin@line}%
      \stepcounter{pinlineno}%
    }
    \closein\pin@file
    \expandafter\endgroup
    \scantokens\expandafter{\pin@accu}%
  }{%
    \errmessage{File `#3' doesn't exist!}%
  }%
}
\makeatother

%%% In line Comments
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\cil}[1]{\todo[inline, color=green!40]{#1}}

%%% Margin Comments
\newcommand{\cm}[1]{\todo[color=green!40]{\begin{adjustwidth*}{1mm}{}#1\end{adjustwidth*}}}
\newcommand{\cmb}[1]{\todo[color=blue!40]{\begin{adjustwidth*}{1mm}{}\lstinline|#1|\end{adjustwidth*}}}

%%% Ref counters
\newcounter{myhypothesis}
\newcommand\hypothesis[1]
{
  \refstepcounter{myhypothesis}\label{hyp_#1}%
}
%
%\twolabels{4}{a}
%I have two labels: \ref{hyp_4} and \ref{hyp_a}.

%%%%%%%%%%%% Formatting %%%%%%%%%%%%

%%% Item Spacing
\newlength{\wideitemsep}
\setlength{\wideitemsep}{\itemsep}
\addtolength{\wideitemsep}{-10pt}
\let\olditem\item
\renewcommand{\item}{\setlength{\itemsep}{\wideitemsep}\olditem}

%%% Margins
\addtolength{\marginparwidth}{6cm}
\addtolength{\paperwidth}{\marginparwidth-1cm}

%%%%%%%%%%%% Title %%%%%%%%%%%%
\title{Intelligence and Age at First Intercourse: Cause or Confound?}
\shorttitle{Intelligence and AFI}
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}

\abstract{Your abstract here.\\\today\ at \currenttime}

\authornote{This material is based upon work that has been supported by the National Science Foundation Graduate Research Fellowship Program under Grant No. (DGE-1445197) and various means of institutional support from the following universities: University of British Columbia \& Vanderbilt University.}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\maketitle
%\linenumbers%\section{Introduction}\section{ }\vspace{-.8cm}
\cmb{content/introduction.txt}\input{content/introduction.txt}

\cmb{content/current_study.txt}\section{Current Study}\label{current_study}\vspace{-.61cm}\input{content/current_study.txt}

\section{Methods}\cmb{content/methods.txt}\input{content/methods.txt}

\cmb{content/Gen1measure.txt}\input{content/Gen1measure.txt}

\cmb{content/Gen2measure.txt}\input{content/Gen2measure.txt}

\section{Results}\cmb{content/results.txt}\input{content/results.txt}
%\subsection{Between vs. Within Illustration}
%\cmb{content/results_meancomp.txt}\input{content/results_meancomp.txt}

\subsection{Between Family Anlayses}\cmb{content/results_between.txt}\input{content/results_between.txt}

\subsection{Within Family Analyses}
\cmb{content/results_between.txt}\input{content/results_within.txt}

\section{Discussion}

\newpage\section{Tables}\label{appen_tables}

\subsection{Descriptive Statistics}
\begin{longtable}{@{\extracolsep{5pt}}lcccccc}
\caption{Gen1 Summary Statistics by Sibling Status.}\label{table_summary_stats_sibinsampleg1}
\partialinput{2}{10}{table_summary_stats_sibinsampleg1.tex}
\end{longtable}\linebreak

\begin{longtable}{@{\extracolsep{5pt}}lcccccc}
\caption{Gen2 Summary Statistics by Sibling Status.}\label{table_summary_stats_sibinsampleg2}
\partialinput{2}{10}{table_summary_stats_sibinsampleg2.tex}
\end{longtable}\pagebreak
\npnoround
\begin{longtable}{@{\extracolsep{5pt}}lllcc}
\caption{Gen2 AFI by Gender, Race, and GenderxRace.}\label{table_afi_race_gender}
\partialinput{2}{18}{table_summary_stats_AFIRACEGENDER.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 9.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{table_g2_9measurement.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P.Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{table_g2loading_9.tex}
\end{longtable}\pagebreak

\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{24}{table_Mean_Mom_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{24}{table_Mean_Child_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{26}{table_Mean_Joint_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{28}{table_Dif_Mom_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{28}{table_Dif_Child_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{32}{table_Dif_Joint_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}
\pagebreak
\section{Figures}\label{figures}
\noindent\begin{minipage}{\textwidth}
    \centering
    \captionof{figure}{Comparable Between Family and Within Family Effects}
     \includegraphics[width=\textwidth]{plots/betweenewithin.png}
	\vspace*{15mm}
     \captionof{figure}{Unequal Between Family and Within Family Effects}
    \includegraphics[width=\textwidth]{plots/betweennewithin.png}
	%\label{betweennewithin}
\end{minipage}

%\begin{comment}
%\begin{figure}
%  \centering
%    \includegraphics[height=.85\textwidth,angle=270]{G1wrightMap}
%     \caption{Wright Map of Gen1 Cognitive Ability}\label{G1wrightMap}
%\end{figure}
%\end{comment}

\appendix\label{appen}
\section{Age 10.5 Replication}\label{appen10}
\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 10.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{table_g2_10measurement.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{table_g2loading_10.tex}
\end{longtable}\pagebreak
\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{24}{table_Mean_Mom_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{24}{table_Mean_Child_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{26}{table_Mean_Joint_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{28}{table_Dif_Mom_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{28}{table_Dif_Child_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{32}{table_Dif_Joint_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}

\section{Age 11.5 Replication}\label{appen11}

\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 11.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{table_g2_11measurement.tex}
\end{longtable}\pagebreak
\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P.Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{table_g2loading_11.tex}
\end{longtable}\pagebreak
\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{24}{table_Mean_Mom_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{24}{table_Mean_Child_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{26}{table_Mean_Joint_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{28}{table_Dif_Mom_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{28}{table_Dif_Child_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{32}{table_Dif_Joint_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}
%\section{Notes}
%\input{content/NOTES.txt}

\bibliography{AFI}
\end{document}