### 2015-09-29 afidraftRnw.tex

\documentclass[a4paper,man,apacite,natbib,12pt]{apa6}
%%%%%%%%%%%% Packages %%%%%%%%%%%%
%%% Formatting
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{verbatim}
\usepackage[yyyymmdd,hhmmss]{datetime}

%%% Page size and length etc
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{changepage}
\usepackage{lineno} %Line numbering

%%% Lists
\usepackage{enumitem}%resume numbering

%%% Graphics
\usepackage{graphicx}
\usepackage[graphicx]{realboxes}

%%% Animate
\usepackage{animate}
\usepackage{media9}

%%% References
\usepackage{varioref} %references page location
\usepackage{hyperref}
\usepackage[longnamesfirst]{natbib}
\usepackage[natbibapa,nodoi]{apacite}

%%%%%%%%%%%% Formatting %%%%%%%%%%%%
%%% Foot Notes into Endnotes
%\usepackage{endnotes}
%\let\footnote\endnote

%%% Item Spacing
\newlength{\wideitemsep}
\setlength{\wideitemsep}{\itemsep}
\addtolength{\wideitemsep}{-10pt}
\let\olditem\item
\renewcommand{\item}{\setlength{\itemsep}{\wideitemsep}\olditem}

%%% Margins
\addtolength{\marginparwidth}{6cm}
\addtolength{\paperwidth}{\marginparwidth-1cm}

%%% Nicer Looking Square Root
\let\oldsqrt\sqrt
% it defines the new \sqrt in terms of the old one
\def\sqrt{\mathpalette\DHLhksqrt}
\def\DHLhksqrt#1#2{%
\setbox0=\hbox{$#1\oldsqrt{#2\,}$}\dimen0=\ht0
\advance\dimen0-0.2\ht0
\setbox2=\hbox{\vrule height\ht0 depth -\dimen0}%
{\box0\lower0.4pt\box2}}

%%%%%%%%%%%% New Commands %%%%%%%%%%%%

%%% Short Hand
\usepackage{xspace}
\newcommand{\R}{\textup{\textrm{R}}\xspace}
\newcommand{\meb}{{\bf Garrison, S. M.}\xspace}
\newcommand{\et}{\textit{et al}.\xspace}
\newcommand{\eg}{\textit{e.g},\xspace}

%%% Remove DOI %UNCOMMENT ALL to toggle between DOI and URL
\usepackage{etoolbox}
\usepackage{environ}

\newtoggle{bibdoi}
\newtoggle{biburl}
\makeatletter

\undef{\APACrefURL}
\undef{\endAPACrefURL}
\undef{\APACrefDOI}
\undef{\endAPACrefDOI}

\long\def\collect@url#1{\global\def\bib@url{#1}}
\long\def\collect@doi#1{\global\def\bib@doi{#1}}
\newenvironment{APACrefURL}{\global\toggletrue{biburl}\Collect@Body\collect@url}{\unskip\unskip}
\newenvironment{APACrefDOI}{\global\toggletrue{bibdoi}\Collect@Body\collect@doi}{}

%\AtBeginEnvironment{thebibliography}{
% \pretocmd{\PrintBackRefs}{%
%  \iftoggle{bibdoi}
%    {\iftoggle{biburl}{\unskip\unskip doi:\bib@doi}{}}
%    {\iftoggle{biburl}{Retrieved from\bib@url}{}}
%  \togglefalse{bibdoi}\togglefalse{biburl}%
%  }{}{}
%}

%%% Partial Input
\usepackage[ascii]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\makeatletter
\newread\pin@file
\newcounter{pinlineno}
\newcommand\pin@accu{}
\newcommand\pin@ext{pintmp}
% inputs #3, selecting only lines #1 to #2 (inclusive)
\newcommand*\partialinput [3] {%
  \IfFileExists{#3}{%
    \openin\pin@file #3
    % skip lines 1 to #1 (exclusive)
    \setcounter{pinlineno}{1}
    \@whilenum\value{pinlineno}<#1 \do{%
      \read\pin@file to\pin@line
      \stepcounter{pinlineno}%
    }
    % prepare reading lines #1 to #2 inclusive
    \addtocounter{pinlineno}{-1}
    \let\pin@accu\empty
    \begingroup
    \endlinechar\newlinechar
    \@whilenum\value{pinlineno}<#2 \do{%
      % use safe catcodes provided by e-TeX's \readline
      \readline\pin@file to\pin@line
      \edef\pin@accu{\pin@accu\pin@line}%
      \stepcounter{pinlineno}%
    }
    \closein\pin@file
    \expandafter\endgroup
    \scantokens\expandafter{\pin@accu}%
  }{%
    \errmessage{File `#3' doesn't exist!}%
  }%
}
\makeatother

%%% In line Comments
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\cil}[1]{\todo[inline, color=green!40]{#1}}

%%% Margin Comments
\newcommand{\cm}[1]{\todo[color=green!40]{\begin{adjustwidth*}{1mm}{}#1\end{adjustwidth*}}}
\newcommand{\cmb}[1]{\todo[color=blue!40]{\begin{adjustwidth*}{1mm}{}\lstinline|#1|\end{adjustwidth*}}}

%%% Hypothesis counters
\newcounter{myhypothesis}
\newcommand\hypothesis[1]
{
  \refstepcounter{myhypothesis}\label{hyp_#1}%
}

%%%%%%%%%%%% Title %%%%%%%%%%%%
\title{Intelligence and Age at First Intercourse: Cause or Confound?}
\shorttitle{Intelligence and AFI}
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}

\abstract{Your abstract here.\\\today\ at \currenttime}

\authornote{This material is based upon work that has been supported by [INSERT JOE'S GRANTS HERE] and the National Science Foundation Graduate Research Fellowship Program under Grant No. (DGE-1445197) and various means of institutional support from the following universities: University of British Columbia \& Vanderbilt University.}

\begin{document}
\maketitle
%\linenumbers%\section{Introduction}\section{ }\vspace{-.8cm}
\cmb{content/introduction.txt}\input{content/introduction.txt}

\cmb{content/current_study.txt}\section{Current Study}\label{current_study}\vspace{-.61cm}\input{content/current_study.txt}

\section{Methods}\cmb{content/methods.txt}\input{content/methods.txt}