### 2015-10-29 R/afidraftRnw.tex

\documentclass[a4paper,man,apacite,natbib,12pt]{apa6}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
%%%%%%%%%%%% Packages %%%%%%%%%%%%
%%% Formatting
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{verbatim}
\usepackage[yyyymmdd,hhmmss]{datetime}

%%% Page size and length etc
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{changepage}
\usepackage{lineno} %Line numbering

%%% Lists
\usepackage{enumitem}%resume numbering

%%% Graphics
\usepackage{graphicx}
\usepackage[graphicx]{realboxes}

%%% Animate
\usepackage{animate}
\usepackage{media9}

%%% References
\usepackage{varioref} %references page location
\usepackage{hyperref}
\usepackage[longnamesfirst]{natbib}
\usepackage[natbibapa,nodoi]{apacite}

\documentclass[a4paper,man,apacite,natbib,12pt]{apa6}
%%%%%%%%%%%% Packages %%%%%%%%%%%%
%%%%%% Formating %%%%%%
%% Font
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[utf8x,ascii]{inputenc} %% Additional Characters
%% Math
\usepackage{amsmath}
%% Date
\usepackage[yyyymmdd,hhmmss]{datetime}
%% Lists
\usepackage{enumitem}%resume numbering

%%%%%% Layout %%%%%%
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{changepage}
\usepackage{xspace}

%%%%%% References %%%%%%
\usepackage{varioref} %references page location
\usepackage{hyperref}
\usepackage[longnamesfirst]{natbib}
\usepackage[natbibapa,nodoi]{apacite}

%%%%%% Graphics %%%%%%
\usepackage{graphicx}
\usepackage[graphicx]{realboxes}
%%% Animate
\usepackage{animate}
\usepackage{media9}

%%%%%% R %%%%%%
\usepackage{listings}
\usepackage{inconsolata}
\usepackage{verbatim}
\usepackage{etoolbox}

%%%%%%%%%%%% New Commands %%%%%%%%%%%%
%% Short Hand
\newcommand{\R}{\textup{\textrm{R}}\xspace}
\newcommand{\meb}{{\bf Garrison, S. M.}\xspace}
\newcommand{\et}{\textit{et al}.\xspace}
\newcommand{\eg}{\textit{e.g},\xspace}

%%% Remove DOI %UNCOMMENT ALL to toggle between DOI and URL
\usepackage{environ}
\newtoggle{bibdoi}
%\newtoggle{biburl}
\makeatletter

%\undef{\APACrefURL}
%\undef{\endAPACrefURL}
\undef{\APACrefDOI}
\undef{\endAPACrefDOI}

%\long\def\collect@url#1{\global\def\bib@url{#1}}
\long\def\collect@doi#1{\global\def\bib@doi{#1}}
%\newenvironment{APACrefURL}{\global\toggletrue{biburl}\Collect@Body\collect@url}{\unskip\unskip}
\newenvironment{APACrefDOI}{\global\toggletrue{bibdoi}\Collect@Body\collect@doi}{}

%\AtBeginEnvironment{thebibliography}{
% \pretocmd{\PrintBackRefs}{%
%  \iftoggle{bibdoi}
%    {\iftoggle{biburl}{\unskip\unskip doi:\bib@doi}{}}
%    {\iftoggle{biburl}{Retrieved from\bib@url}{}}
%  \togglefalse{bibdoi}\togglefalse{biburl}%
%  }{}{}
%}
%%% Partial Input
\makeatletter
\newread\pin@file
\newcounter{pinlineno}
\newcommand\pin@accu{}
\newcommand\pin@ext{pintmp}
% inputs #3, selecting only lines #1 to #2 (inclusive)
\newcommand*\partialinput [3] {%
  \IfFileExists{#3}{%
    \openin\pin@file #3
    % skip lines 1 to #1 (exclusive)
    \setcounter{pinlineno}{1}
    \@whilenum\value{pinlineno}<#1 \do{%
      \read\pin@file to\pin@line
      \stepcounter{pinlineno}%
    }
    % prepare reading lines #1 to #2 inclusive
    \addtocounter{pinlineno}{-1}
    \let\pin@accu\empty
    \begingroup
    \endlinechar\newlinechar
    \@whilenum\value{pinlineno}<#2 \do{%
      % use safe catcodes provided by e-TeX's \readline
      \readline\pin@file to\pin@line
      \edef\pin@accu{\pin@accu\pin@line}%
      \stepcounter{pinlineno}%
    }
    \closein\pin@file
    \expandafter\endgroup
    \scantokens\expandafter{\pin@accu}%
  }{%
    \errmessage{File `#3' doesn't exist!}%
  }%
}
\makeatother

%%% Single Line Filler Text
\makeatletter
\newcounter{fillersentencecounter}
\setcounter{fillersentencecounter}{1}
\newread\pin@file
\newcounter{pinlineno}
\newcommand\pin@accu{}
\newcommand\pin@ext{pintmp}

\newcommand*\fillersentence [2] {%
  \IfFileExists{#2}{%
    \openin\pin@file #2

    \setcounter{pinlineno}{1}
    \@whilenum\value{pinlineno}<\value{fillersentencecounter} \do{%
      \read\pin@file to\pin@line
      \stepcounter{pinlineno}%
    }
    \addtocounter{fillersentencecounter}{#1}
    \addtocounter{fillersentencecounter}{-1}
    \addtocounter{pinlineno}{-1}
    \let\pin@accu\empty
    \begingroup
    \endlinechar\newlinechar
    \@whilenum\value{pinlineno}<\value{fillersentencecounter} \do{%
      \readline\pin@file to\pin@line
      \edef\pin@accu{\pin@accu\pin@line}%
      \stepcounter{pinlineno}%
    }
    \addtocounter{fillersentencecounter}{1}
    \closein\pin@file
    \expandafter\endgroup
    \scantokens\expandafter{\pin@accu}%
  }{%
    \errmessage{File `#2' doesn't exist!}%
  }%
}
\makeatother
%%% Nicer Looking Square Root
\let\oldsqrt\sqrt
% it defines the new \sqrt in terms of the old one
\def\sqrt{\mathpalette\DHLhksqrt}
\def\DHLhksqrt#1#2{%
\setbox0=\hbox{$#1\oldsqrt{#2\,}$}\dimen0=\ht0
\advance\dimen0-0.2\ht0
\setbox2=\hbox{\vrule height\ht0 depth -\dimen0}%
{\box0\lower0.4pt\box2}}

%%% In line Comments
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\cil}[1]{\todo[inline, color=green!40]{#1}}

%%% Margin Comments
\newcommand{\cm}[1]{\todo[color=green!40]{\begin{adjustwidth*}{1mm}{}#1\end{adjustwidth*}}}
\newcommand{\cmb}[1]{\todo[color=blue!40]{\begin{adjustwidth*}{1mm}{}\lstinline|#1|\end{adjustwidth*}}}

%%% Hypothesis counters
\newcounter{myhypothesis}
\newcommand\hypothesis[1]
{
  \refstepcounter{myhypothesis}\label{hyp_#1}%
}

%%%%%%%%%%%% Formatting %%%%%%%%%%%%
%%% Foot Notes into Endnotes
%\usepackage{endnotes}
%\let\footnote\endnote

%%% Item Spacing
\newlength{\wideitemsep}
\setlength{\wideitemsep}{\itemsep}
\addtolength{\wideitemsep}{-10pt}
\let\olditem\item
\renewcommand{\item}{\setlength{\itemsep}{\wideitemsep}\olditem}

%%% Margins
\addtolength{\marginparwidth}{6cm}
\addtolength{\paperwidth}{\marginparwidth-1cm}

%%%%%% Style
%%% R Output
%-----------------
%Define the colors you want to use
\definecolor{keywordcolor}{rgb}{0,0.6,0.6}
\definecolor{delimcolor}{rgb}{0.461,0.039,0.102}
\definecolor{Rcommentcolor}{rgb}{0.101,0.043,0.432}
\definecolor{Black}{rgb}{0,0,0}
%-----------------
\lstdefinestyle{Rsettings}{
  language=R,
  basicstyle=\ttfamily\color{Black},
  breaklines=true,
  showstringspaces=true,
  keywords={if, else, function, theFunction, tmp},
  otherkeywords={},
  commentstyle=\itshape\color{Rcommentcolor},
  keywordstyle=\color{keywordcolor},
  moredelim=[s][\color{delimcolor}]{"}{"},
}

%%%%%%%%%%%% Title %%%%%%%%%%%%
\title{Intelligence and Age at First Intercourse: Cause or Confound?}
\shorttitle{Intelligence and AFI}
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}

\abstract{Your abstract here.\\\today\ at \currenttime}

\authornote{This material is based upon work that has been supported by National Institute of Health under Grant No. (R01-HD065865) and the National Science Foundation Graduate Research Fellowship Program under Grant No. (DGE-1445197) and various means of institutional support from the following universities: University of British Columbia, University of Oklahoma \& Vanderbilt University.}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\maketitle
%\linenumbers%\section{Introduction}\section{ }\vspace{-.8cm}
\input{content/introduction.txt}

\section{Current Study}\label{current_study}\vspace{-.61cm}\input{content/current_study.txt}

\section{Methods}\input{content/methods.txt}

\input{content/Gen1measure.txt}

\input{content/Gen2measure.txt}

\input{content/reliabilitya.txt}
\begin{knitrout}\footnotesize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{sigmax} \hlkwb{=} \hlkwd{sqrt}\hlstd{(}\hlnum{1.1881}\hlstd{)}
\hlstd{sigmay} \hlkwb{=} \hlstd{sigmax}
\hlstd{rxx} \hlkwb{=} \hlnum{.87}
\hlstd{ryy} \hlkwb{=} \hlstd{rxx}
\hlstd{rxy}\hlkwb{=}\hlnum{.67}
\hlstd{(sigmax}\hlopt{*}\hlstd{sigmax}\hlopt{*}\hlstd{rxx} \hlopt{+}\hlstd{sigmay}\hlopt{*}\hlstd{sigmay}\hlopt{*}\hlstd{ryy} \hlopt{-}\hlnum{2}\hlopt{*}\hlstd{rxy}\hlopt{*}\hlstd{sigmax}\hlopt{*}\hlstd{sigmay)} \hlopt{/}
  \hlstd{(sigmax}\hlopt{*}\hlstd{sigmax} \hlopt{+} \hlstd{sigmay}\hlopt{*}\hlstd{sigmay} \hlopt{-}\hlnum{2}\hlopt{*}\hlstd{rxy}\hlopt{*}\hlstd{sigmax}\hlopt{*}\hlstd{sigmay)}
\end{alltt}
\begin{verbatim}
## [1] 0.606
\end{verbatim}
\end{kframe}
\end{knitrout}
Accordingly, we can substitute the following values into the above equation to calcualte the difference score reliability for Gen2 AFI.
where, \begin{itemize}
\item $\sigma_{x}^{2}$ and $\sigma_{y}^{2}$  are 4.41 (from Table \vref{table_summary_stats_sibinsampleg2}) %the variance of Gen2 AFI, 
\item $\rho_{xx'}$ and $\rho_{yy'}$ are .76 (from Table \vref{table_measurement_trt_g1afi}), 
\item $\rho_{xy}$ is the AFI correlation between first born cousins is .099.\end{itemize}
\begin{knitrout}\footnotesize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{sigmax} \hlkwb{=} \hlkwd{sqrt}\hlstd{(}\hlnum{4.41}\hlstd{)}
\hlstd{sigmay} \hlkwb{=} \hlstd{sigmax}
\hlstd{rxx} \hlkwb{=} \hlnum{.76}
\hlstd{ryy} \hlkwb{=} \hlstd{rxx}
\hlstd{rxy}\hlkwb{=}\hlnum{.099}
\hlstd{(sigmax}\hlopt{*}\hlstd{sigmax}\hlopt{*}\hlstd{rxx} \hlopt{+}\hlstd{sigmay}\hlopt{*}\hlstd{sigmay}\hlopt{*}\hlstd{ryy} \hlopt{-}\hlnum{2}\hlopt{*}\hlstd{rxy}\hlopt{*}\hlstd{sigmax}\hlopt{*}\hlstd{sigmay)} \hlopt{/}
  \hlstd{(sigmax}\hlopt{*}\hlstd{sigmax} \hlopt{+} \hlstd{sigmay}\hlopt{*}\hlstd{sigmay} \hlopt{-}\hlnum{2}\hlopt{*}\hlstd{rxy}\hlopt{*}\hlstd{sigmax}\hlopt{*}\hlstd{sigmay)}
\end{alltt}
\begin{verbatim}
## [1] 0.734
\end{verbatim}
\end{kframe}
\end{knitrout}
\input{content/reliabilityb.txt}
\section{Results}\input{content/results.txt}
%\subsection{Between vs. Within Illustration}

\subsection{Between Family Anlayses}\input{content/results_between.txt}

\subsection{Within Family Analyses}
\input{content/results_within.txt}
\section{Discussion}
\input{content/discussion.txt}

\newpage\section{Tables}\label{appen_tables}

\subsection{Descriptive Statistics}
\begin{longtable}{@{\extracolsep{5pt}}lcccccc}
\caption{Gen1 Summary Statistics by Sibling Status.}\label{table_summary_stats_sibinsampleg1}
\partialinput{2}{10}{content/table/table_summary_stats_sibinsampleg1.tex}
\end{longtable}\linebreak

\begin{longtable}{@{\extracolsep{5pt}}lcccccc}
\caption{Gen2 Summary Statistics by Sibling Status.}\label{table_summary_stats_sibinsampleg2}
\partialinput{2}{10}{content/table/table_summary_stats_sibinsampleg2.tex}
\end{longtable}\pagebreak
\npnoround
\begin{longtable}{@{\extracolsep{5pt}}lllcc}
\caption{Gen2 AFI by Gender, Race, and Gender by Race.}\label{table_afi_race_gender}
\partialinput{2}{18}{content/table/table_summary_stats_AFIRACEGENDER.tex}
\end{longtable}
\pagebreak
\subsection{Measurement}
\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 9.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{content/table/table_g2_9measurement.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P.Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{content/table/table_g2loading_9.tex}
\end{longtable}\pagebreak

\subsection{Difference Scores}
\begin{minipage}{\linewidth}
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen1 Self-Reported AFI Across 1983-1985 }\label{table_measurement_trt_g1afi}
\partialinput{6}{12}{content/table/table_ttafireliable_z}
\end{longtable}
\end{minipage}
\pagebreak
\begin{minipage}{\linewidth}
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen1 Sibling Differences in Self-Reported AFI across 1983-1985 (Standardized by Gender)}\label{table_measurement_diffafireliable_z}
\partialinput{6}{12}{content/table/table_diffafireliable_z}
\end{longtable}
\end{minipage}
\noindent\begin{minipage}{\textwidth}
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen2 Aggregated Intelligence across Ages 9.5, 10.5, 11.5 }\label{table_measurement_trt_g2int}
\partialinput{6}{12}{content/table/table_ttintreliable_z}
\end{longtable}
\end{minipage}

\pagebreak
\begin{minipage}{\linewidth}
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen2 First Born Cousin Differences in Intelligence across Ages 9.5-11.5}\label{table_measurement_diffintreliable_z}
\partialinput{6}{12}{content/table/table_diffintreliable_z}
\end{longtable}
\end{minipage}
\pagebreak

\begin{minipage}{\linewidth}
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen2 First Female Cousin Differences in Intelligence across Ages 9.5-11.5}\label{table_measurement_diffintreliable_g}
\partialinput{6}{12}{content/table/table_diffintreliable_g}
\end{longtable}
\end{minipage}
\pagebreak

\begin{minipage}{\linewidth}
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen2 First Male Cousin Differences in Intelligence across Ages 9.5-11.5}\label{table_measurement_diffintreliable_b}
\partialinput{6}{12}{content/table/table_diffintreliable_b}
\end{longtable}
\end{minipage}
\pagebreak
\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{24}{content/table/table_Mean_Mom_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{24}{content/table/table_Mean_Child_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_9}
\partialinput{5}{26}{content/table/table_Mean_Joint_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{28}{content/table/table_Dif_Mom_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{28}{content/table/table_Dif_Child_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_9}
\partialinput{5}{32}{content/table/table_Dif_Joint_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}
\pagebreak
\section{Figures}\label{figures}
\noindent\begin{minipage}{\textwidth}
    \centering
    \captionof{figure}{Comparable Between Family and Within Family Effects}
     \includegraphics[width=\textwidth]{plots/betweenewithin.png}
	\vspace*{15mm}
     \captionof{figure}{Unequal Between Family and Within Family Effects}
    \includegraphics[width=\textwidth]{plots/betweennewithin.png}
	%\label{betweennewithin}
\end{minipage}

%\begin{comment}
%\begin{figure}
%  \centering
%    \includegraphics[height=.85\textwidth,angle=270]{G1wrightMap}
%     \caption{Wright Map of Gen1 Cognitive Ability}\label{G1wrightMap}
%\end{figure}
%\end{comment}

\appendix\label{appen}
%\section{Endnotes}
%\theendnotes

\section{Age 10.5 Replication}\label{appen10}
\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 10.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{content/table/table_g2_10measurement.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{content/table/table_g2loading_10.tex}
\end{longtable}\pagebreak

\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{24}{content/table/table_Mean_Mom_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{24}{content/table/table_Mean_Child_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_10}
\partialinput{5}{26}{content/table/table_Mean_Joint_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{28}{content/table/table_Dif_Mom_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{28}{content/table/table_Dif_Child_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_10}
\partialinput{5}{32}{content/table/table_Dif_Joint_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}

\section{Age 11.5 Replication}\label{appen11}

\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 11.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{content/table/table_g2_11measurement.tex}
\end{longtable}\pagebreak
\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P.Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{content/table/table_g2loading_11.tex}
\end{longtable}\pagebreak
\subsection{Between Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{24}{content/table/table_Mean_Mom_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen2 Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{24}{content/table/table_Mean_Child_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Mean Gen2 AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_11}
\partialinput{5}{26}{content/table/table_Mean_Joint_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak
\subsection{Within Family Analyses}
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen1 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{28}{content/table/table_Dif_Mom_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Gen2 Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{28}{content/table/table_Dif_Child_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}\pagebreak

\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Dif Gen2 AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_11}
\partialinput{5}{32}{content/table/table_Dif_Joint_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}
%\section{Notes}
%\input{content/NOTES.txt}

\bibliography{Projects-AFI}
\end{document}