### 2015-11-09 Manuscript/Manuscript.Rnw
changes

\documentclass[a4paper,man,apacite,natbib,12pt,longtable]{apa6}
\usepackage{../Common/style/Mason}
\usepackage{listings}
\usepackage{inconsolata}

%%%%%%%%%%%% Title %%%%%%%%%%%%
\title{Intelligence and Age at First Intercourse: Cause or Confound?}
\shorttitle{Intelligence and AFI}
%
% Authors and Affiliations
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}
%
% Author Note
\authornote{\input{../Common/content/title/authornote.txt}}
%
% Abstract
\abstract{Last compile was \today\ at \currenttime\\
\input{../Common/content/title/abstract_empreport.txt}}
%
% Key Words
\keywords{Age at first intercourse; Cross-sectional data; Intelligence; Quasi-Experimental}
%
\begin{document}
\maketitle
%
%%%%%%%%%%%% R %%%%%%%%%%%%
<<sweave,echo=FALSE>>=
  options(width=60)
  options(scipen=999)
  options(digits=3)
  options(width=60)

  listing <- function(x, options){
    paste("\\begin{lstlisting}[style=Rsettings]\n",
      x, "\\end{lstlisting}\n", sep = "")}
  knit_hooks$set(source=listing, output=listing)
    knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed = TRUE)})
@
<<setup, include=FALSE,cache=FALSE>>=
# Libraries
## Install Most Current Version
library(devtools)#;dev_mode(on=T)
#install_github("LiveOak/NlsyLinks")
#install_github("harrelfe/Hmisc");dev_mode(on=F)
## Attach Remaining
library(corrplot)
library(ggplot2);library(ggExtra)
library(Hmisc)
library(MplusAutomation)
library(NlsyLinks)
library(plyr);library(psych)
library(stargazer)
library(tables)
library(xtable)
#
# File Structure
file_drive= "C:"
##file_drive="E:"
directory = paste0(file_drive,"/Dropbox/Work/Research/Projects/NLSY/AFI")
#
# Functions
## source("https://raw.githubusercontent.com/smasongarrison/Functions/master/Mason_Functions.R")
## cor_1star
### Prints out clean correlation matrix. If you want to replace the upper triangle with the sample size, toggle include.n, whose default is FALSE.
cor_1star <- function(x,digit=3,sig=.05,include.n=FALSE){ 
  require(Hmisc) 
  x <- as.matrix(x) 
  R <- rcorr(x)$r 
  p <- rcorr(x)$P

  ## define notions for significance levels; spacing is important.
  mystars <- ifelse(p < sig, "* ", " ")

  ## truncate the matrix that holds the correlations to two decimal
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), digit))[,-1]

  ## build a new matrix that includes the correlations with their appropriate stars 
  Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x)) 
  diag(Rnew) <- paste(diag(R), " ", sep="") 
  rownames(Rnew) <- colnames(x) 
  colnames(Rnew) <- paste(colnames(x), "", sep="")

  Rnew <- as.matrix(Rnew)
  if(!include.n){
    ## remove upper triangle
    Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
    Rnew <- as.data.frame(Rnew) 
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
  }
  if(include.n){
    #replace upper triangle with sample size
    require(psych)
    ct<-corr.test(x)
    ct<-as.matrix(ct$n)
    Rnew[upper.tri(Rnew, diag = TRUE)] <- ct[upper.tri(ct, diag = TRUE)]
    Rnew <- as.data.frame(Rnew) 
    Rnew <- cbind(Rnew[1:length(Rnew)])#-1])
  }

  return(Rnew) 
}
##
## ROUNDING FUNCTIONS
### Function rounds to three digits and defaults to na.rm = TRUE
### Cor
Cor <- function(x,digit=3, use="pairwise.complete.obs") {
  round(cor(x, use = use),digit)}

### Max
Max <- function(x,digit=3,na.rm = TRUE) {
  round(max(x, na.rm=na.rm),digit)}

### Mean
Mean <- function(x,digit=3,na.rm = TRUE) {
  round(mean(x, na.rm=na.rm),digit)}

### Median
Median <- function(x,digit=3,na.rm = TRUE) {
  round(median(x, na.rm=na.rm),digit)}

### Min
Min <- function(x,digit=3,na.rm = TRUE) {
  round(min(x, na.rm=na.rm),digit)}

### RowMedians
RowMedians <-function(x,digit=3,na.rm = TRUE) {
  round(rowMedians(as.matrix(x), na.rm=na.rm),digit)}

### Sd
Sd <- function(x,digit=3,na.rm = TRUE) {
  round(sd(x, na.rm=na.rm),digit)}
## Reverse Substring
substrRight <- function(x, n, fromend=0, end=nchar(x)-fromend){
  substr(x, end-n+1,end)}
@
<<setup_Data_9,include=FALSE,cache=TRUE>>=
# Import Data for age 9
## First Born Child
LinkedMDAN_discord_AFI_z_9 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_z_910811.csv"))
LinkedMDAN_discord_AFI_z_9$Gendermean<-as.factor(LinkedMDAN_discord_AFI_z_9$Gendermean)
##
## First Born Girl
LinkedMDAN_discord_AFI_g_9 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_girls_910811.csv"))
##
## First Born Boy
LinkedMDAN_discord_AFI_b_9 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_boys_910811.csv"))
#
# Models for age 9
## First Born Child
data=LinkedMDAN_discord_AFI_z_9
model.meankids_9<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.int_9<-lm(Int_Mean ~ INTMOMmean, data = data)
model.meanmoms_9<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.diffkids_9 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms_9 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids_9 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean +INTMOMmean + Intdiff+  INTMOMdiff, data = data)
model.meanmomskids_9<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffmomsint_9 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff + Gendermean + INTMOMmean:Gendermean, data = data)
##
## First Born Girl
data=LinkedMDAN_discord_AFI_g_9
model.meankids.g_9<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.g_9<-lm(AFIKIDmean ~ INTMOMmean, data = data)
nested<-data
nested$Int_Mean[is.na(nested$Intdiff)] <- NA
model.diffkids.g_9<- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffkids.g.full_9 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = nested)
model.diffkids.g.nested_9 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean , data = nested)
anova_9<-anova(model.diffkids.g.nested_9,model.diffkids.g.full_9)
model.diffmoms.g_9<- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.g_9 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
model.meanmomskids.g_9<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
##
## First Born Boy
data=LinkedMDAN_discord_AFI_b_9
model.meankids.b_9<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.b_9<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.meanmomskids.b_9<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffkids.b_9 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms.b_9 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.b_9 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
@
<<setup_Data_10,include=FALSE,cache=TRUE>>=
# Import Data for age 10
## First Born Child
LinkedMDAN_discord_AFI_z_10 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_z_1011912.csv"))
LinkedMDAN_discord_AFI_z_10$Gendermean<-as.factor(LinkedMDAN_discord_AFI_z_10$Gendermean)
##
## First Born Girl
LinkedMDAN_discord_AFI_g_10 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_girls_1011912.csv"))
##
## First Born Boy
LinkedMDAN_discord_AFI_b_10 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_boys_1011912.csv"))
#
# Models for age 10
## First Born Child
data=LinkedMDAN_discord_AFI_z_10
model.meankids_10<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.int_10<-lm(Int_Mean ~ INTMOMmean, data = data)
model.meanmoms_10<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.diffkids_10 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms_10 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids_10 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean +INTMOMmean + Intdiff+  INTMOMdiff, data = data)
model.meanmomskids_10<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffmomsint_10 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff + Gendermean + INTMOMmean:Gendermean, data = data)
##
## First Born Girl
data=LinkedMDAN_discord_AFI_g_10
model.meankids.g_10<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.g_10<-lm(AFIKIDmean ~ INTMOMmean, data = data)
nested<-data
nested$Int_Mean[is.na(nested$Intdiff)] <- NA
model.diffkids.g_10<- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffkids.g.full_10 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = nested)
model.diffkids.g.nested_10 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean , data = nested)
anova_10<-anova(model.diffkids.g.nested_10,model.diffkids.g.full_10)
model.diffmoms.g_10<- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.g_10 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
model.meanmomskids.g_10<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
##
## First Born Boy
data=LinkedMDAN_discord_AFI_b_10
model.meankids.b_10<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.b_10<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.meanmomskids.b_10<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffkids.b_10 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms.b_10 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.b_10 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
@
<<setup_Data_11,include=FALSE,cache=TRUE>>=
# Import Data for age 10
## First Born Child
LinkedMDAN_discord_AFI_z_11 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_z_11121013.csv"))
LinkedMDAN_discord_AFI_z_11$Gendermean<-as.factor(LinkedMDAN_discord_AFI_z_11$Gendermean)
##
## First Born Girl
LinkedMDAN_discord_AFI_g_11 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_girls_11121013.csv"))
##
## First Born Boy
LinkedMDAN_discord_AFI_b_11 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_boys_11121013.csv"))
#
# Models for age 11
## First Born Child
data=LinkedMDAN_discord_AFI_z_11
model.meankids_11<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.int_11<-lm(Int_Mean ~ INTMOMmean, data = data)
model.meanmoms_11<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.diffkids_11 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms_11 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids_11 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean +INTMOMmean + Intdiff+  INTMOMdiff, data = data)
model.meanmomskids_11<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffmomsint_11 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff + Gendermean + INTMOMmean:Gendermean, data = data)
##
## First Born Girl
data=LinkedMDAN_discord_AFI_g_11
model.meankids.g_11<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.g_11<-lm(AFIKIDmean ~ INTMOMmean, data = data)
nested<-data
nested$Int_Mean[is.na(nested$Intdiff)] <- NA
model.diffkids.g_11<- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffkids.g.full_11 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = nested)
model.diffkids.g.nested_11 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean , data = nested)
anova_11<-anova(model.diffkids.g.nested_11,model.diffkids.g.full_11)
model.diffmoms.g_11<- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.g_11 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
model.meanmomskids.g_11<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
##
## First Born Boy
data=LinkedMDAN_discord_AFI_b_11
model.meankids.b_11<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.b_11<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.meanmomskids.b_11<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffkids.b_11 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms.b_11 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.b_11 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
@
<<setup_summarystats,include=FALSE,cache=TRUE>>=
# Import Data
## Determine First Born
birthweight <- read.table(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/SASGen1Kids/birthweight.dat",sep=""), quote="\"", stringsAsFactors=FALSE)
names(birthweight)<-c('CASEID','MCASEID','RACE','SEX_G2','YRB','WGT')
birthweight$WGT[birthweight$WGT < 0] <- NA
birthweight$firstborn<-0
birthweight$firstborn<-substrRight(birthweight$CASEID,2)
birthweight_2<-birthweight
birthweight$born[birthweight$firstborn!="01"]<-0
birthweight$born[birthweight$firstborn=="01"]<-1
birthweight_fb <- birthweight[ which(birthweight$born==1), ]
##
## Gen 2 Demographics
new_data <- read.table(paste0(directory,'/Common/content/data/gen2demographics.dat'), sep=' ')
names(new_data) <- c('C0000100','C0000200','C0005300','C0005400')
### Handle missing values
  new_data[new_data == -1] = NA  # Refused 
  new_data[new_data == -2] = NA  # Don't know 
  new_data[new_data == -3] = NA  # Invalid missing 
  new_data[new_data == -7] = NA  # Missing 
### If there are values not categorized they will be represented as NA
vallabels = function(data) {
  data$C0005300 <- factor(data$C0005300, levels=c(1.0,2.0,3.0), labels=c("Hispanic","Black","Non-Black, Non-Hispanic"))
  data$C0005400 <- factor(data$C0005400, levels=c(1.0,2.0), labels=c("Male","Female"))
  return(data)}
varlabels <- c("ID CODE OF CHILD",
    "ID CODE OF MOTHER OF CHILD",
    "RACE OF CHILD (FROM MOTHERS SCREENER 79)",
    "SEX OF CHILD")
qnames = function(data) {
  names(data) <- c("CASEID","MCASEID","RACE","CSEX")
  return(data)
}
categories_g2 <- qnames(vallabels(new_data))
momintel <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/SASGen1/momintel.csv",sep=""), header=FALSE)
names(momintel)<-c("MCASEID","Intel_G1","Intel_G1_SE","HID")
##
## Gen2 Mean Intelligence
kidintel<-read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/kidsintl/cleaned_CFA_G_around_910811.csv",sep=""), header=FALSE)
names(kidintel)<-c("CASEID","ASEID","Intel_G2","Intel_G2_SE","MCASEID")
##
## Gen1 AFI
Gen1AFI_exportconcor <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/Gen1AFI_exportconcor_910811.csv",sep=""))
Gen1AFI_exportconcor <- Gen1AFI_exportconcor[2:length(Gen1AFI_exportconcor[,1]),]
names(Gen1AFI_exportconcor)<-c("MCASEID","AFI_G1")
##
## Gen2 Mean AFI
Gen2AFI_exportconcor <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/Gen2AFI_exportconcor_910811.csv",sep=""))
names(Gen2AFI_exportconcor)<-c("CASEID","AFI_G2")
##
# Merge
Gen1AFIConcor <- merge(momintel,Gen1AFI_exportconcor,by="MCASEID",all=T)
Gen2AFIConcor <- merge(kidintel,Gen2AFI_exportconcor,by="CASEID",all=T)
Gen2AFIConcor_2 <- merge(categories_g2,Gen2AFI_exportconcor,by="CASEID",all=T)
@
<<setup_Reliability,include=FALSE,cache=TRUE>>=
# Reliability 
## AFI G1
###Import Data
new_data <- read.table(paste0(directory,"/Common/content/data/afi_reliability.dat"), sep=' ')
names(new_data) <- c('R0000100','R0000149','R0214700','R0214800','R0988000','R1020700','R1312200','R1379600','R1379700','R1379800','R1691600','R1760000','R1760100','R1760200')

#### Handle missing values
  new_data[new_data == -1] = NA  # Refused 
  new_data[new_data == -2] = NA  # Don't know 
  new_data[new_data == -3] = NA  # Invalid missing 
  new_data[new_data == -4] = NA  # Valid missing 
  new_data[new_data == -5] = NA  # Non-interview

#### Assign Values
vallabels = function(data) {
  data$R0214700 <- factor(data$R0214700, levels=c(1.0,2.0,3.0), labels=c("Hispanic","Black","Non-Black, Non-Hispanic"))
  data$R0214800 <- factor(data$R0214800, levels=c(1.0,2.0), labels=c("Male","Female"))
  return(data)
}
#### Assign Labels
varlabels <- c(    "ID# (1-12686) 79",
    "HH ID # 79",
    "RACL/ETHNIC COHORT /SCRNR 79",
    "SEX OF R 79",
    "MALE AGE 1ST HAD INTERCOURSE 83",
    "F AGE 1ST HAD SEXUAL INTERCOURSE 83",
    "MALE AGE 1ST HAD INTERCOURSE 84",
    "F AGE 1ST HAD SEXUAL INTERCOURSE 84",
    "F USE B-DT YR 1ST INTERCOURSE 84",
    "F MO SEXUAL INTERCOURSE 1ST TIME 84",
    "MALE AGE 1ST HAD INTERCOURSE 85",
    "F AGE 1ST HAD SEXUAL INTERCOURSE 85",
    "F USE B-DT YR/1ST INTERCOURSE 85",
    "F MO SEXUAL INTERCOURSE 1ST TIME 85"
)
qnames = function(data) {
  names(data) <- c("CASEID","HHID","RACE","SEX","MAFI_1983","FAFI_1983","MAFI_1984","FAFI_1984","FAFI_YEAR_1984","FAFI_MON_1984","MAFI_1985","FAFI_1985","FAFI_YEAR_1985","FAFI_MON_1985")
  return(data)
}
categories <- qnames(vallabels(new_data))
###
### Trim Variables
testretest<-categories[,c(1,2,4)]
testretest$AFI_83<-as.numeric(NA);testretest$AFI_84<-as.numeric(NA);testretest$AFI_85<-as.numeric(NA);testretest$AFI_83z<-as.numeric(NA);testretest$AFI_84z<-as.numeric(NA);testretest$AFI_85z<-as.numeric(NA)
### Rescale by Gender
for(i in 1:length(categories[,1])){
  if(testretest$SEX[i]=="Male"){
    testretest$AFI_83[i]<-as.numeric(categories$MAFI_1983[i])
    testretest$AFI_84[i]<-as.numeric(categories$MAFI_1984[i])
    testretest$AFI_85[i]<-as.numeric(categories$MAFI_1985[i])
    testretest$AFI_83z[i]<-(as.numeric(categories$MAFI_1983[i])-15.88)/2.15
    testretest$AFI_84z[i]<-(as.numeric(categories$MAFI_1984[i])-15.88)/2.15
    testretest$AFI_85z[i]<-(as.numeric(categories$MAFI_1985[i])-15.88)/2.15
  }
    if(testretest$SEX[i]=="Female"){
    testretest$AFI_83[i]<-as.numeric(categories$FAFI_1983[i])
    testretest$AFI_84[i]<-as.numeric(categories$FAFI_1984[i])
    testretest$AFI_85[i]<-as.numeric(categories$FAFI_1985[i])
    testretest$AFI_83z[i]<-(as.numeric(categories$FAFI_1983[i])-16.47)/1.99
    testretest$AFI_84z[i]<-(as.numeric(categories$FAFI_1984[i])-16.47)/1.99
    testretest$AFI_85z[i]<-(as.numeric(categories$FAFI_1985[i])-16.47)/1.99
    }
}
### Link Siblings
#### Name Prep
names(momintel)<-c("CASEID","Intel_G1","Intel_G1_SE","HID") 
testretest<-merge(testretest,momintel[c(1,2)],by="CASEID", all=T)

#### Link
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen1Housemates")
dsOutcomes <- testretest
dsOutcomes$SubjectTag <-dsOutcomes$CASEID*100
dsSingle <- CreatePairLinksSingleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c("AFI_83","AFI_84","AFI_85","AFI_83z","AFI_84z","AFI_85z"))
dsD <- CreatePairLinksDoubleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c(
"AFI_83","AFI_84","AFI_85","AFI_83z","AFI_84z","AFI_85z","Intel_G1","SEX"
))
dsD_nm<-dsD[dsD$SEX_S1=="FEMALE"]
dsSingle$DIF_AFI_83z<-dsSingle$AFI_83z_S1-dsSingle$AFI_83z_S2
dsSingle$DIF_AFI_84z<-dsSingle$AFI_84z_S1-dsSingle$AFI_84z_S2
dsSingle$DIF_AFI_85z<-dsSingle$AFI_85z_S1-dsSingle$AFI_85z_S2

### Correlation Matrices
matrix_g1<-data.frame(dsD$Intel_G1_S1,dsD$Intel_G1_S2)
matrix_afi_diff<-data.frame(dsSingle$DIF_AFI_83z,dsSingle$DIF_AFI_84z,dsSingle$DIF_AFI_85z)
names(matrix_afi_diff)<-c("1983","1984","1985")
matrix_tt_z<-data.frame(dsD$AFI_83z_S1,dsD$AFI_84z_S1,dsD$AFI_85z_S1)
names(matrix_tt_z)<-c("1983","1984","1985")

capture.output(xtable(cor_1star(matrix_afi_diff,digit=2,sig=.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_diffafireliable_z.tex"))
               capture.output(xtable(cor_1star(matrix_tt_z,digit=2,sig=.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_ttafireliable_z.tex"))
capture.output(xtable(cor_1star(matrix_g1,digit=2,sig=.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_sibintreliable_z.tex"))

## AFI g2
### Import Data
total_data <- read.csv(paste0(directory,"/Common/content/data/gen2demographics.csv"), stringsAsFactors=FALSE)
dsOutcomes <- data.frame(Gen2AFIConcor,total_data$CSEX)
dsOutcomes$AFI_G2z<-as.numeric(NA)

### Rescale by Gender
mean_s1<-mean(dsOutcomes$AFI_G2[dsOutcomes$total_data.CSEX==1],na.rm=TRUE)
mean_s2<-mean(dsOutcomes$AFI_G2[dsOutcomes$total_data.CSEX==2],na.rm=TRUE)
sd_s1<- sd(dsOutcomes$AFI_G2[dsOutcomes$total_data.CSEX==1],na.rm=TRUE)
sd_s2<-sd(dsOutcomes$AFI_G2[dsOutcomes$total_data.CSEX==2],na.rm=TRUE)
dsOutcomes<-merge(dsOutcomes[c(1,2,6,8)],birthweight_fb, by="CASEID")
for(i in 1:nrow(dsOutcomes)){
  if(dsOutcomes$SEX_G2[i]==2&!is.na(dsOutcomes$SEX_G2[i])
){
    dsOutcomes$AFI_G2z[i]<- (dsOutcomes$AFI_G2[i]-mean_s2)/sd_s2
  }
  if(dsOutcomes$SEX_G2[i]==1&!is.na(dsOutcomes$SEX_G2[i])
){
    dsOutcomes$AFI_G2z[i]<- (dsOutcomes$AFI_G2[i]-mean_s1)/sd_s1
  }
}

### Link
dsOutcomes$SubjectTag<-dsOutcomes$MCASEID*100
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen1Housemates")
dsSinglez <- CreatePairLinksSingleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c(
"AFI_G2","AFI_G2z","SEX_G2")
)
dsDz <- CreatePairLinksDoubleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c(
"AFI_G2","AFI_G2z","SEX_G2")
)
### Export Matrix
matrix_afi_cousincor<-data.frame(dsSinglez$AFI_G2z_S1,dsSinglez$AFI_G2z_S2)
capture.output(xtable(cor_1star(matrix_afi_cousincor,2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_sibafig2_z.tex"))
Cor(matrix_afi_cousincor)

## Gen2 Mean Intelligence
### Test-Retest
#### Import
G_G2_9.5 <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/kidsintl/cleaned_CFA_G_around_910811.csv",sep=""), header=FALSE)
names(G_G2_9.5)<-c("CASEID","ASEID","G_G2_9.5","G_SE_G2_9.5","MCASEID")
G_G2_10.5 <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/kidsintl/cleaned_CFA_G_around_1011912.csv",sep=""), header=FALSE)
names(G_G2_10.5)<-c("CASEID","ASEID","G_G2_10.5","G_SE_G2_10.5","MCASEID")
G_G2_11.5 <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/kidsintl/cleaned_CFA_G_around_11121013.csv",sep=""), header=FALSE)
names(G_G2_11.5)<-c("CASEID","ASEID","G_G2_11.5","G_SE_G2_11.5","MCASEID")
#### Merge
temp<-merge(G_G2_9.5[c(1,3)],G_G2_10.5[c(1,3)],by="CASEID",all=T)
G_G2<-merge(temp,G_G2_11.5[c(1,3)],by="CASEID",all=T)
#### Export Matrix Test Retest
capture.output(xtable(cor_1star(G_G2[2:4],2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_ttintreliable_z.tex"))

### First Born Child
#### Import Data
LinkedMDAN_discord_AFI_z_9_m<-LinkedMDAN_discord_AFI_z_9[,c(2,56,66,68,70,72,90)]
names(LinkedMDAN_discord_AFI_z_9_m) <-c("ID" ,"AFIMOMdiff_9", "INTMOMdiff_9","AFMKIDdiff_9","AFIKIDdiff_9","CSMOMdiff_9",  "Intdiff_9" ) 
LinkedMDAN_discord_AFI_z_10_m<-LinkedMDAN_discord_AFI_z_10[,c(2,56,66,68,70,72,90)]
names(LinkedMDAN_discord_AFI_z_10_m) <-c("ID" ,"AFIMOMdiff_10", "INTMOMdiff_10","AFMKIDdiff_10","AFIKIDdiff_10","CSMOMdiff_10",  "Intdiff_10" ) 
LinkedMDAN_discord_AFI_z_11_m<-LinkedMDAN_discord_AFI_z_11[,c(2,56,66,68,70,72,90)]
names(LinkedMDAN_discord_AFI_z_11_m) <-c("ID" ,"AFIMOMdiff_11", "INTMOMdiff_11","AFMKIDdiff_11","AFIKIDdiff_11","CSMOMdiff_11",  "Intdiff_11" )
#### Merge Data
diffscorereliable_z <- merge(LinkedMDAN_discord_AFI_z_9_m,LinkedMDAN_discord_AFI_z_10_m,by="ID",all=T)
diffscorereliable_z <- merge(diffscorereliable_z,LinkedMDAN_discord_AFI_z_11_m,by="ID",all=T)
matrix_z<-data.frame(diffscorereliable_z$Intdiff_9,diffscorereliable_z$Intdiff_10,diffscorereliable_z$Intdiff_11)
names(matrix_z)<-c("Age 9.5","Age 10.5","Age 11.5")

### First Born Girls
#### Import Data
LinkedMDAN_discord_AFI_g_9_m<-LinkedMDAN_discord_AFI_g_9[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_g_9_m) <-c("ID" ,"AFIMOMdiff_9", "INTMOMdiff_9","AFMKIDdiff_9","AFIKIDdiff_9","CSMOMdiff_9",  "Intdiff_9" ) 
LinkedMDAN_discord_AFI_g_10_m<-LinkedMDAN_discord_AFI_g_10[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_g_10_m) <-c("ID" ,"AFIMOMdiff_10", "INTMOMdiff_10","AFMKIDdiff_10","AFIKIDdiff_10","CSMOMdiff_10",  "Intdiff_10" ) 
LinkedMDAN_discord_AFI_g_11_m<-LinkedMDAN_discord_AFI_g_11[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_g_11_m) <-c("ID" ,"AFIMOMdiff_11", "INTMOMdiff_11","AFMKIDdiff_11","AFIKIDdiff_11","CSMOMdiff_11",  "Intdiff_11" ) 
#### Merge Data
diffscorereliable_g <- merge(LinkedMDAN_discord_AFI_g_9_m,LinkedMDAN_discord_AFI_g_10_m,by="ID",all=T)
diffscorereliable_g <- merge(diffscorereliable_g,LinkedMDAN_discord_AFI_g_11_m,by="ID",all=T)
matrix_g<-data.frame(diffscorereliable_g$Intdiff_9,diffscorereliable_g$Intdiff_10,diffscorereliable_g$Intdiff_11)
names(matrix_g)<-c("Age 9.5","Age 10.5","Age 11.5")

### First Born Boys
#### Import Data
LinkedMDAN_discord_AFI_b_9_m<-LinkedMDAN_discord_AFI_b_9[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_b_9_m) <-c("ID" ,"AFIMOMdiff_9", "INTMOMdiff_9","AFMKIDdiff_9","AFIKIDdiff_9","CSMOMdiff_9",  "Intdiff_9" ) 
LinkedMDAN_discord_AFI_b_10_m<-LinkedMDAN_discord_AFI_b_10[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_b_10_m) <-c("ID" ,"AFIMOMdiff_10", "INTMOMdiff_10","AFMKIDdiff_10","AFIKIDdiff_10","CSMOMdiff_10",  "Intdiff_10" ) 
LinkedMDAN_discord_AFI_b_11_m<-LinkedMDAN_discord_AFI_b_11[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_b_11_m) <-c("ID" ,"AFIMOMdiff_11", "INTMOMdiff_11","AFMKIDdiff_11","AFIKIDdiff_11","CSMOMdiff_11",  "Intdiff_11" ) 
#### Merge
diffscorereliable_b <- merge(LinkedMDAN_discord_AFI_b_9_m,LinkedMDAN_discord_AFI_b_10_m,by="ID",all=T)
diffscorereliable_b <- merge(diffscorereliable_b,LinkedMDAN_discord_AFI_b_11_m,by="ID",all=T)
matrix_b<-data.frame(diffscorereliable_b$Intdiff_9,diffscorereliable_b$Intdiff_10,diffscorereliable_b$Intdiff_11)
names(matrix_b)<-c("Age 9.5","Age 10.5","Age 11.5")

### Export Cousin Cor
capture.output(xtable(cor_1star(matrix_b,2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_diffintreliable_b.tex"))
capture.output(xtable(cor_1star(matrix_g,2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_diffintreliable_g.tex"))
capture.output(xtable(cor_1star(matrix_z,2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_diffintreliable_z.tex"))
@
<<measurement,include=FALSE,cache=TRUE>>=
# G1
## mplus
### Run
#### runModels(directory =paste0(directory,"/Common/content/mplusautomation/MplusG1"))
ModelSummaries_Mplus<-extractModelSummaries(target = paste0(directory,"/Common/content/mplusautomation/MplusG1"))
### Clean Up
ModelSummaries_Mplus[,4]<-NULL;ModelSummaries_Mplus[,3]<-NULL;ModelSummaries_Mplus[,2]<-NULL;ModelSummaries_Mplus[,1]<-NULL;ModelSummaries_Mplus[,23]<-NULL
### Export
capture.output(stargazer(ModelSummaries_Mplus,summary=FALSE,flip = TRUE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_gen1measurement.tex"))

# G2
## Mplus
### Run
#### runModels(directory =paste0(directory,"/Common/content/mplusautomation/MplusG2"))
g2loading_9 <- read.csv(paste0(directory,"/Common/content/data/g2loading_9.csv"), stringsAsFactors=FALSE)
g2loading_10 <- read.csv(paste0(directory,"/Common/content/data/g2loading_10.csv"), stringsAsFactors=FALSE)
g2loading_11 <- read.csv(paste0(directory,"/Common/content/data/g2loading_11.csv"), stringsAsFactors=FALSE)
### Extract
ModelSummaries_Mplusg2_9<-extractModelSummaries(target = paste0(directory,"/Common/content/mplusautomation/MplusG2/g_around_910811.out"))
ModelSummaries_Mplusg2_10<-extractModelSummaries(target = paste0(directory,"/Common/content/mplusautomation/MplusG2/g_around_1011912.out"))
ModelSummaries_Mplusg2_11<-extractModelSummaries(target = paste0(directory,"/Common/content/mplusautomation/MplusG2/g_around_11121013.out"))
### Clean
ModelSummaries_Mplusg2_9[,4]<-NULL;ModelSummaries_Mplusg2_9[,3]<-NULL;ModelSummaries_Mplusg2_9[,2]<-NULL;ModelSummaries_Mplusg2_9[,1]<-NULL;ModelSummaries_Mplusg2_9[,26]<-NULL;ModelSummaries_Mplusg2_9[,7]<-NULL;ModelSummaries_Mplusg2_9[,7]<-NULL;ModelSummaries_Mplusg2_9[,7]<-NULL

ModelSummaries_Mplusg2_10[,4]<-NULL;ModelSummaries_Mplusg2_10[,3]<-NULL;ModelSummaries_Mplusg2_10[,2]<-NULL;ModelSummaries_Mplusg2_10[,1]<-NULL;ModelSummaries_Mplusg2_10[,26]<-NULL;ModelSummaries_Mplusg2_10[,7]<-NULL;ModelSummaries_Mplusg2_10[,7]<-NULL;ModelSummaries_Mplusg2_10[,7]<-NULL

ModelSummaries_Mplusg2_11[,4]<-NULL;ModelSummaries_Mplusg2_11[,3]<-NULL;ModelSummaries_Mplusg2_11[,2]<-NULL;ModelSummaries_Mplusg2_11[,1]<-NULL;ModelSummaries_Mplusg2_11[,26]<-NULL;ModelSummaries_Mplusg2_11[,7]<-NULL;ModelSummaries_Mplusg2_11[,7]<-NULL;ModelSummaries_Mplusg2_11[,7]<-NULL
### Export
#### Age 9
capture.output(stargazer(ModelSummaries_Mplusg2_9,summary=FALSE,flip = TRUE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2_9measurement.tex"))
capture.output(stargazer(g2loading_9,summary=FALSE,flip = FALSE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2loading_9.tex"))

#### Age 10
capture.output(stargazer(ModelSummaries_Mplusg2_10,summary=FALSE,flip = TRUE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2_10measurement.tex"))
capture.output(stargazer(g2loading_10,summary=FALSE,flip = FALSE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2loading_10.tex"))

#### Age 11
capture.output(stargazer(ModelSummaries_Mplusg2_11,summary=FALSE,flip = TRUE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2_11measurement.tex"))
capture.output(stargazer(g2loading_11,summary=FALSE,flip = FALSE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2loading_11.tex"))
@
<<results_regression,echo=FALSE, comment=NA, include=FALSE,cache=TRUE>>=
# Stargazer Plot defaults
font.size="small"
no.space=TRUE
style="demography"
column.sep.width = "1pt"
header = FALSE
column.labels = c("Mixed", "Daughters", "Sons")
model.numbers=FALSE
float=FALSE
ci =TRUE
single.row = TRUE
# 9.5
capture.output(stargazer(model.meanmoms_9,model.meanmoms.g_9,model.meanmoms.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_9.tex"))

capture.output(stargazer(model.meankids_9,model.meankids.g_9,model.meankids.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_9.tex"))

capture.output(stargazer(model.meanmomskids_9,model.meanmomskids.g_9,model.meanmomskids.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel","Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_9.tex"))

capture.output(stargazer(model.diffmoms_9,model.diffmoms.g_9,model.diffmoms.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen1 Mean Intel", "Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_9.tex"))

capture.output(stargazer(model.diffkids_9,model.diffkids.g_9,model.diffkids.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel", "Gen2 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_9.tex"))

capture.output(stargazer(model.diffmomskids_9,model.diffmomskids.g_9,model.diffmomskids.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel","Gen1 Mean Intel","Gen2 Dif Intel","Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_9.tex"))

# 10.5
capture.output(stargazer(model.meanmoms_10,model.meanmoms.g_10,model.meanmoms.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_10.tex"))

capture.output(stargazer(model.meankids_10,model.meankids.g_10,model.meankids.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_10.tex"))

capture.output(stargazer(model.meanmomskids_10,model.meanmomskids.g_10,model.meanmomskids.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel","Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_10.tex"))

capture.output(stargazer(model.diffmoms_10,model.diffmoms.g_10,model.diffmoms.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen1 Mean Intel", "Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_10.tex"))

capture.output(stargazer(model.diffkids_10,model.diffkids.g_10,model.diffkids.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel", "Gen2 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_10.tex"))

capture.output(stargazer(model.diffmomskids_10,model.diffmomskids.g_10,model.diffmomskids.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel","Gen1 Mean Intel","Gen2 Dif Intel","Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_10.tex"))

# 11.5
capture.output(stargazer(model.meanmoms_11,model.meanmoms.g_11,model.meanmoms.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_11.tex"))

capture.output(stargazer(model.meankids_11,model.meankids.g_11,model.meankids.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_11.tex"))

capture.output(stargazer(model.meanmomskids_11,model.meanmomskids.g_11,model.meanmomskids.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel","Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_11.tex"))

capture.output(stargazer(model.diffmoms_11,model.diffmoms.g_11,model.diffmoms.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen1 Mean Intel", "Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_11.tex"))

capture.output(stargazer(model.diffkids_11,model.diffkids.g_11,model.diffkids.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel", "Gen2 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_11.tex"))

capture.output(stargazer(model.diffmomskids_11,model.diffmomskids.g_11,model.diffmomskids.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel","Gen1 Mean Intel","Gen2 Dif Intel","Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_11.tex"))

# Summary Stats by Race and Gender
capture.output(latex(
  tabular((Heading()*( 1 + RowFactor(CSEX)+ RowFactor(RACE) + RowFactor(CSEX)*RowFactor(RACE))~(Heading("AFI")*AFI_G2*(Mean + Sd))), 
          data = Gen2AFIConcor_2),
  numeric.dollar=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_AFIRACEGENDER.tex"))
@
<<setup_tables,include=FALSE,cache=TRUE>>=
# Summary Stats G1
## Manage data
### Import
df <- merge(Gen1AFIConcor,birthweight_fb,by="MCASEID",all=T)
nlsyextract <- read.csv(paste0(directory,"/Common/content/data/nlsyextract.csv"))
###
### Clean
nlsyextract$IDC1[nlsyextract$IDC1 <= 0  ] <- 0
nlsyextract$IDC1[nlsyextract$IDC1 > 0  ] <- 1
nlsyextract$SiblinginSample <-nlsyextract$IDC1
nlsyextract$IDC1 <- NULL
nlsyextract[nlsyextract< 0  ] <- NA
nlsyextract$month<-nlsyextract$Q1.3_A.M 
nlsyextract$month<- paste("0",nlsyextract$Q1.3_A.M,sep="")
nlsyextract$month[nlsyextract$month == 010 ] <- 10
nlsyextract$month[nlsyextract$month == 011 ] <- 11
nlsyextract$month[nlsyextract$month == 012 ] <- 12
nlsyextract$Age<-as.numeric(as.Date(paste(df$YRB,"-06-15",sep=""), format='%Y-%m-%d')-as.Date(paste("19",nlsyextract$Q1.3_A.Y,"-",nlsyextract$month,"-15",sep=""), format='%Y-%m-%d'))/365.25
summarystats_g1<-data.frame(nlsyextract$SiblinginSample,nlsyextract$Age,df$Intel_G1,df$AFI_G1,nlsyextract$SAMPLE_SEX)
###
### Sibling in sample?
#### Gen2AFIConcor$MCASEID<-birthweight$MCASEID
#### data=Gen2AFIConcor
#### data$dup<-0
#### for (x in 1:length(data[,1])) {
#### for (y in 1:length(data[,1])) {
#### if (data$MCASEID[x]==data$MCASEID[y]) data$dup[x]= data$dup[x]+1 }}
#### write.table(data, file = paste0(directory,"/Common/content/data/g2sibinsample.csv"), append = FALSE)
Gen2AFIConcor<- read.table(paste0(directory,"/Common/content/data/g2sibinsample.csv"), header=TRUE, quote="\"")
Gen2AFIConcor$SiblinginSample<-1
Gen2AFIConcor$SiblinginSample[Gen2AFIConcor$dup == 1 ]<- 0
summarystats_g2 <- merge(categories_g2,Gen2AFIConcor,by="CASEID",all=T)
summarystats_g2$zIntel_G2<-scale(summarystats_g2$Intel_G2)
summarystats_g2 <- merge(summarystats_g2,birthweight_2,by="CASEID",all=T)
##
## Capture output
capture.output(latex(
  tabular(((Heading("Mother Age at Birth")*nlsyextract.Age + Heading("Standardized AFQT")*df.Intel_G1 + Heading("AFI")*df.AFI_G1 ) ~ Heading("Has Sibling in NLSY?")*(Mean + Sd)*Heading()*( 1 + RowFactor(nlsyextract.SiblinginSample) )), data = summarystats_g1),
  numeric.dollar=TRUE,na.blank=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_sibinsampleg1.tex"))

capture.output(latex(
  tabular(Heading("Year Born")*YRB + Heading("Standardized Ability Score")*zIntel_G2 + Heading("AFI")*AFI_G2 ~ (Heading("Has Sibling in NLSY?")*(Mean + Sd)*Heading()*(1 + RowFactor(SiblinginSample))), data = summarystats_g2),
  numeric.dollar=TRUE,na.blank=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_sibinsampleg2.tex"))
@

%%%%%%%%%%%% Introduction %%%%%%%%%%%%
%\section{Introduction}
\section{ }\vspace{-.8cm}
% Introduce the Major Problem
\input{../Common/content/introduction/introduce_the_problem.txt}\\
\input{../Common/content/introduction/importance_of_problem.txt}

% Relevant Scholarship
%\subsection{Relevant Scholarship}
\input{../Common/content/introduction/relevant_scholarship.txt}\\

% Current Study
\subsection{Current Study}
\input{../Common/content/introduction/current_study.txt}

%%%%%%%%%%%% Method %%%%%%%%%%%%
\section{Method}
% Research Design
\subsection{Research Design}
\input{../Common/content/method/design.txt}
%
% Sample
\subsection{Sample}
\input{../Common/content/method/subjects.txt}
%
% Tetrads
\subsection{Tetrads}
\input{../Common/content/method/tetrad.txt}
%
% Age at First Intercourse
\subsection{Age at First Intercourse}
\input{../Common/content/method/AFI.txt}
%
%%%%%%%%%%%% Measurement %%%%%%%%%%%%
\section{Measures}
%
% Gen1
\subsection{Generation 1}
\input{../Common/content/method/Gen1measures.txt}
%
% Gen2
\subsection{Generation 2}
\input{../Common/content/method/Gen2measures.txt}
%
% Difference Score Reliability
\subsection{Reliability of Difference Scores}
\input{../Common/content/method/diffscores.txt}
<<calculationint,cache=TRUE>>=
sigmax = sqrt(1.1881)
sigmay = sigmax
rxx = .87
ryy = rxx
rxy=.67

(sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) /
  (sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay)
@
<<calculationint2,include=FALSE,cache=TRUE>>=
r.diff.intelligence <- (sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) /
  (sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay)
@
The calculated reliability of Generation 1's differences in AFQT was \Sexpr{r.diff.intelligence}, acceptable, but lower than the empirical correlation we derived for cousin differences.\\

We can also calculate the difference score reliability for Generation 2 AFI, by substituting the following values into equation \ref{equation_Lord1963},
where, \begin{itemize}
\item $\sigma_{x}^{2}$ and $\sigma_{y}^{2}$ are 4.41 (from Table \vref{table_summary_stats_sibinsampleg2}) %the variance of Gen2 Mean AFI, 
\item $\rho_{xx'}$ and $\rho_{yy'}$ are .76 (from Table \vref{table_measurement_trt_g1afi}), 
\item $\rho_{xy}$ is the AFI correlation between first born cousins is .099.(from Figure \vref{plot_between})\end{itemize}
<<calculationafi,cache=TRUE>>=
sigmax = sqrt(4.41)
sigmay = sigmax
rxx = .76
ryy = rxx
rxy=.099
(sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) / 
  (sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay)
@
<<calculationafi2,cache=TRUE,include=FALSE>>=
r.diff.afi<- (sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) / 
  (sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay)
@
 Gen2 Mean AFI difference scores were also reliable (r = \Sexpr{r.diff.afi}) and comparable to Generation 1 sibling differences.
%%%%%%%%%%%% Results %%%%%%%%%%%%
\section{Results}
\input{../Common/content/results/results.txt}
%
% Between Family
\subsection{Between Family}
\input{../Common/content/results/results_between.txt}
%
% Within Family
\subsection{Within Family}
\input{../Common/content/results/results_within.txt}
%
%%%%%%%%%%%% Discussion %%%%%%%%%%%%
\section{Discussion}
\input{../Common/content/discussion/discussion.txt}

%%%%%%%%%%%% Tables %%%%%%%%%%%%
\newpage\section{Tables}
% Descriptive Statistics
\subsection{Descriptive Statistics}
%
%% Sibling in Sample Summary Stats
%%% Gen 1
\begin{longtable}{@{\extracolsep{5pt}}lcccccc}
\caption{Gen1 Summary Statistics by Sibling Status.}\label{table_summary_stats_sibinsampleg1}
\partialinput{2}{10}{../Common/content/tables/table_summary_stats_sibinsampleg1.tex}
\end{longtable}\linebreak
%%%
%%% Gen 2
\begin{longtable}{@{\extracolsep{5pt}}lcccccc}
\caption{Gen2 Summary Statistics by Sibling Status.}\label{table_summary_stats_sibinsampleg2}
\partialinput{2}{10}{../Common/content/tables/table_summary_stats_sibinsampleg2.tex}
\end{longtable}\pagebreak
%%
%% AFI by Gender, Race, and Gender by Race
\npnoround
\begin{longtable}{@{\extracolsep{5pt}}lllcc}
\caption{Gen2 Mean AFI by Gender, Race, and Gender by Race.}\label{table_afi_race_gender}
\partialinput{2}{18}{../Common/content/tables/table_summary_stats_AFIRACEGENDER.tex}
\end{longtable}\pagebreak
%%
% Measurement
\subsection{Measurement}
%%
%% Age 9
%%% Measurement Model
\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 9.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{../Common/content/tables/table_g2_9measurement.tex}
\end{longtable}\pagebreak
%%%
%%% Factor Loadings
\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P.Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{../Common/content/tables/table_g2loading_9.tex}
\end{longtable}\pagebreak
%
% Difference Scores
\subsection{Difference Scores}
%
%% AFI
%%% AFI Test-Retest Gen 1
\begin{minipage}{\linewidth}
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen1 Self-Reported AFI Across 1983-1985 }\label{table_measurement_trt_g1afi}
\partialinput{6}{12}{../Common/content/tables/table_ttafireliable_z.tex}
\end{longtable}
\linebreak
%%%
%%% AFI Sibling Differences
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen1 Sibling Differences in Self-Reported AFI across 1983-1985 (Standardized by Gender)}\label{table_measurement_diffafireliable_z.tex}
\partialinput{6}{12}{../Common/content/tables/table_diffafireliable_z.tex}
\end{longtable}
%%
%% Intelligence
%%% Test-retest Gen 2 intelligence
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen2 Aggregated Intelligence across Ages 9.5, 10.5, 11.5 }\label{table_measurement_trt_g2int}
\partialinput{6}{12}{../Common/content/tables/table_ttintreliable_z.tex}
\end{longtable}\end{minipage}
\pagebreak
\newpage
%%%
%%% Cousin Difference Scores Gen 2 intelligence
%%%% First Born Child
\begin{minipage}{\linewidth}
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen2 First Born Cousin Differences in Intelligence across Ages 9.5-11.5}\label{table_measurement_diffintreliable_z}
\partialinput{6}{12}{../Common/content/tables/table_diffintreliable_z.tex}
\end{longtable}
%%%%
%%%% First Born Daughter
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen2 First Female Cousin Differences in Intelligence across Ages 9.5-11.5}\label{table_measurement_diffintreliable_g}
\partialinput{6}{12}{../Common/content/tables/table_diffintreliable_g.tex}
\end{longtable}
%%%%
%%%% First Born Son
\begin{longtable}{@{\extracolsep{5pt}}rlll} \caption{Correlation of Gen2 First Male Cousin Differences in Intelligence across Ages 9.5-11.5}\label{table_measurement_diffintreliable_b}
\partialinput{6}{12}{../Common/content/tables/table_diffintreliable_b.tex}
\end{longtable}\end{minipage}
\pagebreak
%
% Between Family Analyses
\begin{landscape}
\subsection{Between Family Analyses}
%%
%% Mom -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen1 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{22}{../Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak
%%
%% Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen2 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{22}{../Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak
%%
%% Mom Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{23}{../Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_9.tex}
\end{longtable}\pagebreak
%
% Within Family Analyses
\subsection{Within Family Analyses}
%%
%% Mom -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen1 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{24}{../Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\pagebreak
%%
%% Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen2 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{24}{../Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\pagebreak
%%
%% Mom Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_9}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{26}{../Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_9.tex}
\end{longtable}\end{landscape}

%%%%%%%%%%%% Figures %%%%%%%%%%%%
\section{Figures}
\begin{center}
\captionof{figure}{Smoothed Density Plots of AFI by Sex and Race}
\label{plot_afi_by_race_sex}
<<plot_afi_by_race_sex,fig=TRUE,echo=FALSE,cache=FALSE,echo=FALSE, fig.height=5*.6, fig.width=5,out.width=".8\\paperwidth",warning=FALSE>>=
# Color Blind Palette 
cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7","#000000")
# Data Prep
dataframe<-data.frame(Gen2AFIConcor_2,Gen2AFIConcor$Intel_G2)
names(dataframe)<- c("CASEID","MCASEID","RACE","SEX","AFI","Intel")

## Race by Sex factor
dataframe$RACEbySEX<-"0"
###
### Hispanic
#### Male
dataframe$RACEbySEX[dataframe$RACE=='Hispanic' & dataframe$SEX=='Male']<-"Hispanic Male"
####
#### Female
dataframe$RACEbySEX[dataframe$RACE=='Hispanic'&dataframe$SEX=='Female']<-"Hispanic Female"
###
### Black
#### Male
dataframe$RACEbySEX[dataframe$RACE=='Black'&dataframe$SEX=='Male']<-"Black Male"
####
#### Female
dataframe$RACEbySEX[dataframe$RACE=='Black'&dataframe$SEX=='Female']<-"Black Female"
###
### Non-Black, Non-Hispanic
#### Male
dataframe$RACEbySEX[dataframe$RACE=='Non-Black, Non-Hispanic'&dataframe$SEX=='Male']<-"White Male"
####
#### Female
dataframe$RACEbySEX[dataframe$RACE=='Non-Black, Non-Hispanic'&dataframe$SEX=='Female']<-"White Female"
###
### Unknown Race
dataframe$RACEbySEX[dataframe$RACEbySEX=='0']<-"Unknown"
dataframe$RACE<-revalue(dataframe$RACE, c("Non-Black, Non-Hispanic"="White"))
#summary(as.factor(dataframe$RACEbySEX))

ggplot(dataframe, aes(x=AFI, colour=RACE, group=RACEbySEX, linetype = SEX))+ geom_density(fill=NA,adjust=2)+ theme_bw() + theme(axis.text=element_text(size=10), axis.title=element_text(size=12), plot.title=element_text(size=12)) +  theme(legend.position="right",  legend.justification=c(1, 0.5)) +   xlab("AFI") + ylab("Density") +scale_colour_manual(values= c("#56B4E9", "#009E73", "#CC79A7")) + coord_cartesian(ylim=c(0,0.30), xlim=c(12,25))
@
\end{center}
\pagebreak
%
\begin{center}
<<plot_corplotmatrix_within,fig=TRUE,echo=FALSE,cache=FALSE,fig.height=7*.6, fig.width=7,out.width="1.1\\paperwidth",warning=FALSE>>=
# Croat Cor Matrix Plot for Between Family and Sibling Effects
dataframe<- LinkedMDAN_discord_AFI_z_9
C<-dataframe[,c(8,13,23,24,26,31,41,42)]#,56,57,66,67,70,71
##
## Rename variables
names(C)<-c("Intel_Mom_1","AFI_Mom_1","AFI_Kid_1","Intel_Kid_1","Intel_Mom_2","AFI_Mom_2","AFI_Kid_2","Intel_Kid_2")
##
## Reorder Variables
C<-C[c("Intel_Mom_1","Intel_Mom_2","Intel_Kid_1","Intel_Kid_2","AFI_Mom_1","AFI_Mom_2","AFI_Kid_1","AFI_Kid_2")]
##
## Plot
corrplot(Cor(C), method = "ellipse",type="lower",tl.col="black")
@
\captionof{figure}{Within Family Correlations}
\label{plot_within}
\end{center}
\pagebreak
%
\begin{center}
<<plot_corplotmatrix_between,fig=TRUE,echo=FALSE,cache=FALSE, fig.height=7*.6, fig.width=7,out.width="1.1\\paperwidth",warning=FALSE>>=
# Create Cor Matrix Plot for Between Family and Sibling Effects
dataframe<- LinkedMDAN_discord_AFI_z_9
C<-dataframe[,c(56,57,66,67,70,71,89,90)]
##
## Rename variables
names(C)<-c("AFI_Mom_diff","AFI_Mom_mean","Intel_Mom_diff","Intel_Mom_mean","AFI_Kid_diff","AFI_Kid_mean","Intel_Kid_mean","Intel_Kid_diff")
##
## Reorder Variables
C<-C[c("Intel_Mom_mean","Intel_Mom_diff","Intel_Kid_mean","Intel_Kid_diff","AFI_Mom_mean","AFI_Mom_diff","AFI_Kid_mean","AFI_Kid_diff")]
##
## Plot
corrplot(Cor(C), method = "ellipse",type="lower",tl.col="black")
@
\captionof{figure}{Between Family Correlations}
\label{plot_between}
\end{center}
\pagebreak

\begin{center}
<<plot_spaghetti,fig=TRUE,echo=FALSE,cache=FALSE, fig.height=7, fig.width=7,out.width=".8\\paperwidth",warning=FALSE>>=
# Plot Set up
mytheme <- theme_classic() %+replace% 
  theme(axis.title.x = element_text(face="bold",angle=0), 
        axis.title.y = element_text(face="bold",angle=90)) 
# Data Set Up
dataframe<- LinkedMDAN_discord_AFI_z_9
C<-dataframe[,c(2,8,13,23,24,26,31,41,42)]#,56,57,66,67,70,71
##
## Rename variables
names(C)<-c("ID","Intel_Mom_1","AFI_Mom_1","AFI_Kid_1","Intel_Kid_1","Intel_Mom_2","AFI_Mom_2","AFI_Kid_2","Intel_Kid_2")
##
## Reorder Variables
C<-C[c("ID","Intel_Mom_1","Intel_Mom_2","Intel_Kid_1","Intel_Kid_2","AFI_Mom_1","AFI_Mom_2","AFI_Kid_1","AFI_Kid_2")]

C2<-C
names(C2)<-c("ID","Intel_Mom_2","Intel_Mom_1","Intel_Kid_2","Intel_Kid_1","AFI_Mom_2","AFI_Mom_1","AFI_Kid_2","AFI_Kid_1")
C2<-C2[c("ID","Intel_Mom_1","Intel_Mom_2","Intel_Kid_1","Intel_Kid_2","AFI_Mom_1","AFI_Mom_2","AFI_Kid_1","AFI_Kid_2")]

Stack<-rbind(C,C2)

Stack$slope <- (Stack$AFI_Kid_2-Stack$AFI_Kid_1)/(Stack$Intel_Kid_2-Stack$Intel_Kid_1)
Stack$slopebinary<- as.numeric(NA)
Stack$slopebinary[Stack$slope>0]<- 1
Stack$slopebinary[Stack$slope<0]<- -1
Stack$slopebinary[Stack$slope==0]<- 0
Stack$slopebinary<-factor(Stack$slopebinary, levels=c("1", "0", "-1"), labels=c("Positive", "Zero", "Negative"))
Stack2<-Stack[,c(1,2,4,6,8,11)]

Stack3<-Stack2
### Generate within vs. between slopes

x <- rep(-5:5, length=233)

btwn <- Cor(data.frame(Stack2$AFI_Kid_1,Stack2$Intel_Kid_1))[1,2] * x

within <- Mean(Stack$slope)* x
within2 <- Median(Stack$slope)* x
df <- data.frame(x, y=btwn)

# create a copy of original data, but set y = x
# this data will be used for the one-to-one line
df2 <- data.frame(x, y=within)
df22 <- data.frame(x, y=within2)
# bind original and 'one-to-one data' together
df3 <- rbind.data.frame(df, df2, df22)

# create a grouping variable to separate stat_smoothers based on original and one-to-one data 
df3$grp <- as.factor(rep(1:3, each = nrow(df)))

df3$grp<-factor(df3$grp, levels=c("1", "2", "3"), labels=c("Between Family Slope", "Within Family Slope (mean)", "Within Family Slope (median)"))
df3=df3[1:698,]
# If you rather want to stack the two keys in the legend you can add:
# guide = guide_legend(direction = "vertical")
#...as argument in scale_colour_manual
p1 <- ggplot(data = Stack2, aes(x = Intel_Kid_1, y = AFI_Kid_1, group = ID, color=slopebinary )) +
  mytheme + scale_colour_manual(name = "Within Family \nSlope",values= c("#56B4E9", "#009E73", "#CC79A7")) +
  labs(list(title = "Gen2. Within Family Spaghetti Plot of \nAFI by Intelligence", y = "Gender Standardized AFI", x = "Standardized Intelligence")) + 
  geom_line(size=1,alpha=0.4)  + theme(legend.position="right",  legend.justification=c(1, 0.5)) + coord_cartesian(xlim = c(Min(Stack2$Intel_Kid_1),Max(Stack2$Intel_Kid_1) ),ylim = c(Min(Stack2$AFI_Kid_1),Max(Stack2$AFI_Kid_1)))  +
  stat_smooth(aes(x = df3$x, y = df3$y, group = df3$grp, linetype=df3$grp), method = "lm", se = FALSE, size = 1) + scale_linetype_manual(name = "Slopes",values=c("solid","twodash", "dotted"))

p1 #+
 # geom_smooth(method = "lm", se = FALSE, size = 1)

@
\captionof{figure}{Spaghetti Plot. The mean Within-Family Slope is \Sexpr{Mean(Stack$slope)}; the median Within-Family Slope is \Sexpr{Median(Stack$slope)}}
\label{plot_spaghetti}
\end{center}
%%%%%%%%%%%% References %%%%%%%%%%%%
\bibliography{Projects-AFI}
%
%%%%%%%%%%%% Appendix %%%%%%%%%%%%
\appendix\label{appen}
%
% Age 10.5 Replication
\section{Age 10.5 Replication}\label{appen10}
%% Measurement
%%% Measurement Model
\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 10.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{../Common/content/tables/table_g2_10measurement.tex}
\end{longtable}\pagebreak
%%%
%%% Factor Loadings
\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{../Common/content/tables/table_g2loading_10.tex}
\end{longtable}\pagebreak
%%
%% Between Family
\begin{landscape}
\subsection{Between Family Analyses}
%%%
%%% Mom -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{22}{../Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak
%%%
%%% Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen2 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{22}{../Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak
%%%
%%% Mom Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{23}{../Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_10.tex}
\end{longtable}\pagebreak
%%
%% Within Family
\subsection{Within Family Analyses}
%%%
%%% Mom -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen1 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{24}{../Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}\pagebreak
%%%
%%% Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen2 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{24}{../Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}\pagebreak
%%%
%%% Mom Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Joint Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_10}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{26}{../Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_10.tex}
\end{longtable}
\end{landscape}
%
% Age 11.5 Replication
\section{Age 11.5 Replication}\label{appen11}
%% Measurement
%%% Measurement Model
\begin{longtable}{@{\extracolsep{5pt}}cc} 
\caption{Gen2 Measurement Model.}\label{table_gen2measurement_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & g at Age 11.5 \\ 
\hline \\[-1.8ex] 
\partialinput{12}{34}{../Common/content/tables/table_g2_11measurement.tex}
\end{longtable}\pagebreak
%%%
%%% Factor Loadings
\begin{longtable}{@{\extracolsep{5pt}}cccccc} 
\caption{Gen2 Factor Loadings.}\label{table_g2loading_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & Test & Estimate & S.E. & Est./.S.E. & P.Value \\  
\hline \\[-1.8ex] 
\partialinput{12}{17}{../Common/content/tables/table_g2loading_11.tex}
\end{longtable}\pagebreak
%%
%% Between Family
\begin{landscape}
\subsection{Between Family Analyses}
%%%
%%% Mom -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Gen1 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{22}{../Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak
%%%
%%% Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen2 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{22}{../Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak
%%%
%%% Mom Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Mean Joint Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
\cline{2-4}
\partialinput{10}{23}{../Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_11.tex}
\end{longtable}\pagebreak
%%
%% Within Family
\subsection{Within Family Analyses}
%%%
%%% Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen1 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{24}{../Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}\pagebreak
%%%
%%% Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Gen2 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{24}{../Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}\pagebreak
%%%
%%% Mom Child -> Child
\begin{longtable}{@{\extracolsep{5pt}}lccc} 
\caption{Dif Joint Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_11}
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
\cline{2-4}
\partialinput{10}{26}{../Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_11.tex}
\end{longtable}
\end{landscape}
%
\end{document}