### 2015-12-06 Manuscript/Manuscript.Rnw
changes

\documentclass[a4paper,man,apacite,natbib,12pt,longtable]{apa6}
\usepackage{../Common/style/Mason}
\usepackage{listings}
\usepackage{inconsolata}

%%%%%%%%%%%% Title %%%%%%%%%%%%
\title{Is the Link Between Intelligence and Age at First Intercourse Causal or Correlational? \\ A Cross-Generational Sibling Comparison Design Using the NLSY}
\shorttitle{Intelligence and AFI}
%
% Authors and Affiliations
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}
%
% Author Note
\authornote{\input{../Common/content/title/authornote.txt}}
%
% Abstract
\abstract{Last compile was \today\ at \currenttime\\
\input{../Common/content/title/abstract_empreport.txt}}
%
% Key Words
\keywords{Age at first intercourse; Cross-sectional Data; Discordant Sibling Design; Intelligence; Quasi-Experimental; Siblings}
%
\begin{document}
\maketitle
%
%%%%%%%%%%%% R %%%%%%%%%%%%
<<sweave,echo=FALSE>>=
  options(width=60)
  options(scipen=999)
  options(digits=3)
  options(width=60)

  listing <- function(x, options){
    paste("\\begin{lstlisting}[style=Rsettings]\n",
      x, "\\end{lstlisting}\n", sep = "")}
  knit_hooks$set(source=listing, output=listing)
@
<<setup, include=FALSE>>=
# Libraries
## Install Most Current Version
library(devtools)#;dev_mode(on=T)
#install_github("LiveOak/NlsyLinks")
#install_github("harrelfe/Hmisc");dev_mode(on=F)
## Attach Remaining
library(corrplot)
library(ggplot2);library(ggExtra)
library(Hmisc)
library(MplusAutomation)
library(NlsyLinks)
library(plyr);library(psych)
library(stargazer)
library(tables)
library(xtable)
#
# File Structure
file_drive= "C:"
##file_drive="E:"
directory = paste0(file_drive,"/Dropbox/Work/Research/Projects/Legacy/NLSY/AFI")
#
# Functions
## source("https://raw.githubusercontent.com/smasongarrison/Functions/master/Mason_Functions.R")
## cor_1star
### Prints out clean correlation matrix. If you want to replace the upper triangle with the sample size, toggle include.n, whose default is FALSE.
cor_1star <- function(x,digit=3,sig=.05,include.n=FALSE){ 
  require(Hmisc) 
  x <- as.matrix(x) 
  R <- rcorr(x)$r 
  p <- rcorr(x)$P 
   ## truncate the matrix that holds the correlations to two decimal
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), digit))[,-1] 
  ## build a new matrix that includes the correlations with their appropriate stars 
  Rnew <- R

  if(is.numeric(sig)){
  ## define notions for significance levels; spacing is important.
  mystars <- ifelse(p < sig, "* ", " ")
  ## build a new matrix that includes the correlations with their appropriate stars 
  Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x)) }
  diag(Rnew) <- paste(diag(R), " ", sep="") 
  rownames(Rnew) <- colnames(x) 
  colnames(Rnew) <- paste(colnames(x), "", sep="")

  Rnew <- as.matrix(Rnew)
  if(!include.n){
    ## remove upper triangle
    Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
    Rnew <- as.data.frame(Rnew) 
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
  }
  if(include.n){
    #replace upper triangle with sample size
    require(psych)
    ct<-corr.test(x)
    ct<-as.matrix(ct$n)
    Rnew[upper.tri(Rnew, diag = TRUE)] <- ct[upper.tri(ct, diag = TRUE)]
    Rnew <- as.data.frame(Rnew) 
    Rnew <- cbind(Rnew[1:length(Rnew)])#-1])
  }
  return(Rnew) 
}
##
## ROUNDING FUNCTIONS
### Function rounds to three digits and defaults to na.rm = TRUE
### Cor
Cor <- function(x,digit=3, use="pairwise.complete.obs") {
  round(cor(x, use = use),digit)}

### Max
Max <- function(x,digit=3,na.rm = TRUE) {
  round(max(x, na.rm=na.rm),digit)}

### Mean
Mean <- function(x,digit=3,na.rm = TRUE) {
  round(mean(x, na.rm=na.rm),digit)}

### Median
Median <- function(x,digit=3,na.rm = TRUE) {
  round(median(x, na.rm=na.rm),digit)}

### Min
Min <- function(x,digit=3,na.rm = TRUE) {
  round(min(x, na.rm=na.rm),digit)}

### RowMedians
RowMedians <-function(x,digit=3,na.rm = TRUE) {
  round(rowMedians(as.matrix(x), na.rm=na.rm),digit)}

### Sd
Sd <- function(x,digit=3,na.rm = TRUE) {
  round(sd(x, na.rm=na.rm),digit)}
## Reverse Substring
substrRight <- function(x, n, fromend=0, end=nchar(x)-fromend){
  substr(x, end-n+1,end)}
@
<<setup_Data_9,include=FALSE,cache=TRUE>>=
# Import Data for age 9
## First Born Child
LinkedMDAN_discord_AFI_z_9 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_z_910811.csv"))
LinkedMDAN_discord_AFI_z_9$Gendermean<-as.factor(LinkedMDAN_discord_AFI_z_9$Gendermean)
##
## First Born Girl
LinkedMDAN_discord_AFI_g_9 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_girls_910811.csv"))
##
## First Born Boy
LinkedMDAN_discord_AFI_b_9 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_boys_910811.csv"))
#
# Models for age 9
## First Born Child
data=LinkedMDAN_discord_AFI_z_9
model.meankids_9<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.int_9<-lm(Int_Mean ~ INTMOMmean, data = data)
model.meanmoms_9<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.diffkids_9 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms_9 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids_9 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean +INTMOMmean + Intdiff+  INTMOMdiff, data = data)
model.meanmomskids_9<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffmomsint_9 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff + Gendermean + INTMOMmean:Gendermean, data = data)
##
## First Born Girl
data=LinkedMDAN_discord_AFI_g_9
model.meankids.g_9<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.g_9<-lm(AFIKIDmean ~ INTMOMmean, data = data)
nested<-data
nested$Int_Mean[is.na(nested$Intdiff)] <- NA
model.diffkids.g_9<- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffkids.g.full_9 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = nested)
model.diffkids.g.nested_9 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean , data = nested)
anova_9<-anova(model.diffkids.g.nested_9,model.diffkids.g.full_9)
model.diffmoms.g_9<- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.g_9 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
model.meanmomskids.g_9<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
##
## First Born Boy
data=LinkedMDAN_discord_AFI_b_9
model.meankids.b_9<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.b_9<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.meanmomskids.b_9<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffkids.b_9 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms.b_9 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.b_9 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
@
<<setup_Data_10,include=FALSE,cache=TRUE>>=
# Import Data for age 10
## First Born Child
LinkedMDAN_discord_AFI_z_10 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_z_1011912.csv"))
LinkedMDAN_discord_AFI_z_10$Gendermean<-as.factor(LinkedMDAN_discord_AFI_z_10$Gendermean)
##
## First Born Girl
LinkedMDAN_discord_AFI_g_10 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_girls_1011912.csv"))
##
## First Born Boy
LinkedMDAN_discord_AFI_b_10 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_boys_1011912.csv"))
#
# Models for age 10
## First Born Child
data=LinkedMDAN_discord_AFI_z_10
model.meankids_10<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.int_10<-lm(Int_Mean ~ INTMOMmean, data = data)
model.meanmoms_10<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.diffkids_10 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms_10 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids_10 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean +INTMOMmean + Intdiff+  INTMOMdiff, data = data)
model.meanmomskids_10<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffmomsint_10 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff + Gendermean + INTMOMmean:Gendermean, data = data)
##
## First Born Girl
data=LinkedMDAN_discord_AFI_g_10
model.meankids.g_10<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.g_10<-lm(AFIKIDmean ~ INTMOMmean, data = data)
nested<-data
nested$Int_Mean[is.na(nested$Intdiff)] <- NA
model.diffkids.g_10<- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffkids.g.full_10 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = nested)
model.diffkids.g.nested_10 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean , data = nested)
anova_10<-anova(model.diffkids.g.nested_10,model.diffkids.g.full_10)
model.diffmoms.g_10<- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.g_10 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
model.meanmomskids.g_10<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
##
## First Born Boy
data=LinkedMDAN_discord_AFI_b_10
model.meankids.b_10<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.b_10<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.meanmomskids.b_10<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffkids.b_10 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms.b_10 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.b_10 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
@
<<setup_Data_11,include=FALSE,cache=TRUE>>=
# Import Data for age 10
## First Born Child
LinkedMDAN_discord_AFI_z_11 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_z_11121013.csv"))
LinkedMDAN_discord_AFI_z_11$Gendermean<-as.factor(LinkedMDAN_discord_AFI_z_11$Gendermean)
##
## First Born Girl
LinkedMDAN_discord_AFI_g_11 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_girls_11121013.csv"))
##
## First Born Boy
LinkedMDAN_discord_AFI_b_11 <- read.csv(paste0(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/LinkedMDAN_discord_AFI_boys_11121013.csv"))
#
# Models for age 11
## First Born Child
data=LinkedMDAN_discord_AFI_z_11
model.meankids_11<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.int_11<-lm(Int_Mean ~ INTMOMmean, data = data)
model.meanmoms_11<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.diffkids_11 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms_11 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids_11 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean +INTMOMmean + Intdiff+  INTMOMdiff, data = data)
model.meanmomskids_11<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffmomsint_11 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff + Gendermean + INTMOMmean:Gendermean, data = data)
##
## First Born Girl
data=LinkedMDAN_discord_AFI_g_11
model.meankids.g_11<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.g_11<-lm(AFIKIDmean ~ INTMOMmean, data = data)
nested<-data
nested$Int_Mean[is.na(nested$Intdiff)] <- NA
model.diffkids.g_11<- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffkids.g.full_11 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = nested)
model.diffkids.g.nested_11 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean , data = nested)
anova_11<-anova(model.diffkids.g.nested_11,model.diffkids.g.full_11)
model.diffmoms.g_11<- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.g_11 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
model.meanmomskids.g_11<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
##
## First Born Boy
data=LinkedMDAN_discord_AFI_b_11
model.meankids.b_11<-lm(AFIKIDmean ~ Int_Mean, data = data)
model.meanmoms.b_11<-lm(AFIKIDmean ~ INTMOMmean, data = data)
model.meanmomskids.b_11<-lm(AFIKIDmean ~ INTMOMmean + Int_Mean, data = data)
model.diffkids.b_11 <- lm(AFIKIDdiff ~ AFIKIDmean + Int_Mean + Intdiff, data = data)
model.diffmoms.b_11 <- lm(AFIKIDdiff ~ AFIKIDmean + INTMOMmean + INTMOMdiff, data = data)
model.diffmomskids.b_11 <- lm(AFIKIDdiff ~ AFIKIDmean +Int_Mean + Intdiff+ INTMOMmean + INTMOMdiff, data = data)
@
<<setup_summarystats,include=FALSE,cache=TRUE>>=
# Import Data
## Determine First Born
birthweight <- read.table(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/SASGen1Kids/birthweight.dat",sep=""), quote="\"", stringsAsFactors=FALSE)
names(birthweight)<-c('CASEID','MCASEID','RACE','SEX_G2','YRB','WGT')
birthweight$WGT[birthweight$WGT < 0] <- NA
birthweight$firstborn<-0
birthweight$firstborn<-substrRight(birthweight$CASEID,2)
birthweight_2<-birthweight
birthweight$born[birthweight$firstborn!="01"]<-0
birthweight$born[birthweight$firstborn=="01"]<-1
birthweight_fb <- birthweight[ which(birthweight$born==1), ]
##
## Gen 2 Demographics
new_data <- read.table(paste0(directory,'/Common/content/data/gen2demographics.dat'), sep=' ')
names(new_data) <- c('C0000100','C0000200','C0005300','C0005400')
### Handle missing values
  new_data[new_data == -1] = NA  # Refused 
  new_data[new_data == -2] = NA  # Don't know 
  new_data[new_data == -3] = NA  # Invalid missing 
  new_data[new_data == -7] = NA  # Missing 
### If there are values not categorized they will be represented as NA
vallabels = function(data) {
  data$C0005300 <- factor(data$C0005300, levels=c(1.0,2.0,3.0), labels=c("Hispanic","Black","Non-Black, Non-Hispanic"))
  data$C0005400 <- factor(data$C0005400, levels=c(1.0,2.0), labels=c("Male","Female"))
  return(data)}
varlabels <- c("ID CODE OF CHILD",
    "ID CODE OF MOTHER OF CHILD",
    "RACE OF CHILD (FROM MOTHERS SCREENER 79)",
    "SEX OF CHILD")
qnames = function(data) {
  names(data) <- c("CASEID","MCASEID","RACE","CSEX")
  return(data)
}
categories_g2 <- qnames(vallabels(new_data))
momintel <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/SASGen1/momintel.csv",sep=""), header=FALSE)
names(momintel)<-c("MCASEID","Intel_G1","Intel_G1_SE","HID")
##
## Gen2 Mean Intelligence
kidintel<-read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/kidsintl/cleaned_CFA_G_around_910811.csv",sep=""), header=FALSE)
names(kidintel)<-c("CASEID","ASEID","Intel_G2","Intel_G2_SE","MCASEID")
##
## Gen1 AFI
Gen1AFI_exportconcor <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/CTT/Gen1AFI_exportconcor_910811.csv",sep=""))
Gen1AFI_exportconcor <- Gen1AFI_exportconcor[2:length(Gen1AFI_exportconcor[,1]),]
names(Gen1AFI_exportconcor)<-c("MCASEID","AFI_G1")
##
## Gen2 Mean AFI
Gen2AFI_exportconcor <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/NLSYSAS/Gen2AFI_exportconcor_910811.csv",sep=""))
names(Gen2AFI_exportconcor)<-c("CASEID","AFI_G2")
##
# Merge
Gen1AFIConcor <- merge(momintel,Gen1AFI_exportconcor,by="MCASEID",all=T)
Gen2AFIConcor <- merge(kidintel,Gen2AFI_exportconcor,by="CASEID",all=T)
Gen2AFIConcor_2 <- merge(categories_g2,Gen2AFI_exportconcor,by="CASEID",all=T)
@
<<setup_Reliability,include=FALSE,cache=TRUE>>=
# Reliability 
## AFI G1
###Import Data
new_data <- read.table(paste0(directory,"/Common/content/data/afi_reliability.dat"), sep=' ')
names(new_data) <- c('R0000100','R0000149','R0214700','R0214800','R0988000','R1020700','R1312200','R1379600','R1379700','R1379800','R1691600','R1760000','R1760100','R1760200')

#### Handle missing values
  new_data[new_data == -1] = NA  # Refused 
  new_data[new_data == -2] = NA  # Don't know 
  new_data[new_data == -3] = NA  # Invalid missing 
  new_data[new_data == -4] = NA  # Valid missing 
  new_data[new_data == -5] = NA  # Non-interview

#### Assign Values
vallabels = function(data) {
  data$R0214700 <- factor(data$R0214700, levels=c(1.0,2.0,3.0), labels=c("Hispanic","Black","Non-Black, Non-Hispanic"))
  data$R0214800 <- factor(data$R0214800, levels=c(1.0,2.0), labels=c("Male","Female"))
  return(data)
}
#### Assign Labels
varlabels <- c(    "ID# (1-12686) 79",
    "HH ID # 79",
    "RACL/ETHNIC COHORT /SCRNR 79",
    "SEX OF R 79",
    "MALE AGE 1ST HAD INTERCOURSE 83",
    "F AGE 1ST HAD SEXUAL INTERCOURSE 83",
    "MALE AGE 1ST HAD INTERCOURSE 84",
    "F AGE 1ST HAD SEXUAL INTERCOURSE 84",
    "F USE B-DT YR 1ST INTERCOURSE 84",
    "F MO SEXUAL INTERCOURSE 1ST TIME 84",
    "MALE AGE 1ST HAD INTERCOURSE 85",
    "F AGE 1ST HAD SEXUAL INTERCOURSE 85",
    "F USE B-DT YR/1ST INTERCOURSE 85",
    "F MO SEXUAL INTERCOURSE 1ST TIME 85"
)
qnames = function(data) {
  names(data) <- c("CASEID","HHID","RACE","SEX","MAFI_1983","FAFI_1983","MAFI_1984","FAFI_1984","FAFI_YEAR_1984","FAFI_MON_1984","MAFI_1985","FAFI_1985","FAFI_YEAR_1985","FAFI_MON_1985")
  return(data)
}
categories <- qnames(vallabels(new_data))
###
### Trim Variables
testretest<-categories[,c(1,2,4)]
testretest$AFI_83<-as.numeric(NA);testretest$AFI_84<-as.numeric(NA);testretest$AFI_85<-as.numeric(NA);testretest$AFI_83z<-as.numeric(NA);testretest$AFI_84z<-as.numeric(NA);testretest$AFI_85z<-as.numeric(NA)
### Rescale by Gender
for(i in 1:length(categories[,1])){
  if(testretest$SEX[i]=="Male"){
    testretest$AFI_83[i]<-as.numeric(categories$MAFI_1983[i])
    testretest$AFI_84[i]<-as.numeric(categories$MAFI_1984[i])
    testretest$AFI_85[i]<-as.numeric(categories$MAFI_1985[i])
    testretest$AFI_83z[i]<-(as.numeric(categories$MAFI_1983[i])-15.88)/2.15
    testretest$AFI_84z[i]<-(as.numeric(categories$MAFI_1984[i])-15.88)/2.15
    testretest$AFI_85z[i]<-(as.numeric(categories$MAFI_1985[i])-15.88)/2.15
  }
    if(testretest$SEX[i]=="Female"){
    testretest$AFI_83[i]<-as.numeric(categories$FAFI_1983[i])
    testretest$AFI_84[i]<-as.numeric(categories$FAFI_1984[i])
    testretest$AFI_85[i]<-as.numeric(categories$FAFI_1985[i])
    testretest$AFI_83z[i]<-(as.numeric(categories$FAFI_1983[i])-16.47)/1.99
    testretest$AFI_84z[i]<-(as.numeric(categories$FAFI_1984[i])-16.47)/1.99
    testretest$AFI_85z[i]<-(as.numeric(categories$FAFI_1985[i])-16.47)/1.99
    }
}
### Link Siblings
#### Name Prep
names(momintel)<-c("CASEID","Intel_G1","Intel_G1_SE","HID") 
testretest<-merge(testretest,momintel[c(1,2)],by="CASEID", all=T)

#### Link
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen1Housemates")
dsOutcomes <- testretest
dsOutcomes$SubjectTag <-dsOutcomes$CASEID*100
dsSingle <- CreatePairLinksSingleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c("AFI_83","AFI_84","AFI_85","AFI_83z","AFI_84z","AFI_85z"))
dsD <- CreatePairLinksDoubleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c(
"AFI_83","AFI_84","AFI_85","AFI_83z","AFI_84z","AFI_85z","Intel_G1","SEX"
))
dsD_nm<-dsD[dsD$SEX_S1=="FEMALE"]
dsSingle$DIF_AFI_83z<-dsSingle$AFI_83z_S1-dsSingle$AFI_83z_S2
dsSingle$DIF_AFI_84z<-dsSingle$AFI_84z_S1-dsSingle$AFI_84z_S2
dsSingle$DIF_AFI_85z<-dsSingle$AFI_85z_S1-dsSingle$AFI_85z_S2

### Correlation Matrices
matrix_g1<-data.frame(dsD$Intel_G1_S1,dsD$Intel_G1_S2)
matrix_afi_diff<-data.frame(dsSingle$DIF_AFI_83z,dsSingle$DIF_AFI_84z,dsSingle$DIF_AFI_85z)
names(matrix_afi_diff)<-c("1983","1984","1985")
matrix_tt_z<-data.frame(dsD$AFI_83z_S1,dsD$AFI_84z_S1,dsD$AFI_85z_S1)
names(matrix_tt_z)<-c("1983","1984","1985")

capture.output(xtable(cor_1star(matrix_afi_diff,digit=2,sig=NA,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_diffafireliable_z.tex"))
               capture.output(xtable(cor_1star(matrix_tt_z,digit=2,sig=NA,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_ttafireliable_z.tex"))

capture.output(xtable(cor_1star(matrix_g1,digit=2,sig=NA,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_sibintreliable_z.tex"))

## AFI g2
### Import Data
total_data <- read.csv(paste0(directory,"/Common/content/data/gen2demographics.csv"), stringsAsFactors=FALSE)
dsOutcomes <- data.frame(Gen2AFIConcor,total_data$CSEX)
dsOutcomes$AFI_G2z<-as.numeric(NA)

### Rescale by Gender
mean_s1<-mean(dsOutcomes$AFI_G2[dsOutcomes$total_data.CSEX==1],na.rm=TRUE)
mean_s2<-mean(dsOutcomes$AFI_G2[dsOutcomes$total_data.CSEX==2],na.rm=TRUE)
sd_s1<- sd(dsOutcomes$AFI_G2[dsOutcomes$total_data.CSEX==1],na.rm=TRUE)
sd_s2<-sd(dsOutcomes$AFI_G2[dsOutcomes$total_data.CSEX==2],na.rm=TRUE)
dsOutcomes<-merge(dsOutcomes[c(1,2,6,8)],birthweight_fb, by="CASEID")
for(i in 1:nrow(dsOutcomes)){
  if(dsOutcomes$SEX_G2[i]==2&!is.na(dsOutcomes$SEX_G2[i])
){
    dsOutcomes$AFI_G2z[i]<- (dsOutcomes$AFI_G2[i]-mean_s2)/sd_s2
  }
  if(dsOutcomes$SEX_G2[i]==1&!is.na(dsOutcomes$SEX_G2[i])
){
    dsOutcomes$AFI_G2z[i]<- (dsOutcomes$AFI_G2[i]-mean_s1)/sd_s1
  }
}

### Link
dsOutcomes$SubjectTag<-dsOutcomes$MCASEID*100
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen1Housemates")
dsSinglez <- CreatePairLinksSingleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c(
"AFI_G2","AFI_G2z","SEX_G2")
)
dsDz <- CreatePairLinksDoubleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c(
"AFI_G2","AFI_G2z","SEX_G2")
)
### Export Matrix
matrix_afi_cousincor<-data.frame(dsSinglez$AFI_G2z_S1,dsSinglez$AFI_G2z_S2)
capture.output(xtable(cor_1star(matrix_afi_cousincor,2,NA,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_sibafig2_z.tex"))
Cor(matrix_afi_cousincor)

## Gen2 Mean Intelligence
### Test-Retest
#### Import
G_G2_9.5 <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/kidsintl/cleaned_CFA_G_around_910811.csv",sep=""), header=FALSE)
names(G_G2_9.5)<-c("CASEID","ASEID","G_G2_9.5","G_SE_G2_9.5","MCASEID")
G_G2_10.5 <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/kidsintl/cleaned_CFA_G_around_1011912.csv",sep=""), header=FALSE)
names(G_G2_10.5)<-c("CASEID","ASEID","G_G2_10.5","G_SE_G2_10.5","MCASEID")
G_G2_11.5 <- read.csv(paste(file_drive,"/Dropbox/Software/SAS/MyFolders/79/kidsintl/cleaned_CFA_G_around_11121013.csv",sep=""), header=FALSE)
names(G_G2_11.5)<-c("CASEID","ASEID","G_G2_11.5","G_SE_G2_11.5","MCASEID")
#### Merge
temp<-merge(G_G2_9.5[c(1,3)],G_G2_10.5[c(1,3)],by="CASEID",all=T)
G_G2<-merge(temp,G_G2_11.5[c(1,3)],by="CASEID",all=T)
#### Export Matrix Test Retest
A<-G_G2[2:4]
names(A)<-c("Age 9.5","Age 10.5","Age 11.5")
capture.output(xtable(cor_1star(A,2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_ttintreliable_z.tex"))

### First Born Child
#### Import Data
LinkedMDAN_discord_AFI_z_9_m<-LinkedMDAN_discord_AFI_z_9[,c(2,56,66,68,70,72,90)]
names(LinkedMDAN_discord_AFI_z_9_m) <-c("ID" ,"AFIMOMdiff_9", "INTMOMdiff_9","AFMKIDdiff_9","AFIKIDdiff_9","CSMOMdiff_9",  "Intdiff_9" ) 
LinkedMDAN_discord_AFI_z_10_m<-LinkedMDAN_discord_AFI_z_10[,c(2,56,66,68,70,72,90)]
names(LinkedMDAN_discord_AFI_z_10_m) <-c("ID" ,"AFIMOMdiff_10", "INTMOMdiff_10","AFMKIDdiff_10","AFIKIDdiff_10","CSMOMdiff_10",  "Intdiff_10" ) 
LinkedMDAN_discord_AFI_z_11_m<-LinkedMDAN_discord_AFI_z_11[,c(2,56,66,68,70,72,90)]
names(LinkedMDAN_discord_AFI_z_11_m) <-c("ID" ,"AFIMOMdiff_11", "INTMOMdiff_11","AFMKIDdiff_11","AFIKIDdiff_11","CSMOMdiff_11",  "Intdiff_11" )
#### Merge Data
diffscorereliable_z <- merge(LinkedMDAN_discord_AFI_z_9_m,LinkedMDAN_discord_AFI_z_10_m,by="ID",all=T)
diffscorereliable_z <- merge(diffscorereliable_z,LinkedMDAN_discord_AFI_z_11_m,by="ID",all=T)
matrix_z<-data.frame(diffscorereliable_z$Intdiff_9,diffscorereliable_z$Intdiff_10,diffscorereliable_z$Intdiff_11)
names(matrix_z)<-c("Age 9.5","Age 10.5","Age 11.5")

### First Born Girls
#### Import Data
LinkedMDAN_discord_AFI_g_9_m<-LinkedMDAN_discord_AFI_g_9[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_g_9_m) <-c("ID" ,"AFIMOMdiff_9", "INTMOMdiff_9","AFMKIDdiff_9","AFIKIDdiff_9","CSMOMdiff_9",  "Intdiff_9" ) 
LinkedMDAN_discord_AFI_g_10_m<-LinkedMDAN_discord_AFI_g_10[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_g_10_m) <-c("ID" ,"AFIMOMdiff_10", "INTMOMdiff_10","AFMKIDdiff_10","AFIKIDdiff_10","CSMOMdiff_10",  "Intdiff_10" ) 
LinkedMDAN_discord_AFI_g_11_m<-LinkedMDAN_discord_AFI_g_11[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_g_11_m) <-c("ID" ,"AFIMOMdiff_11", "INTMOMdiff_11","AFMKIDdiff_11","AFIKIDdiff_11","CSMOMdiff_11",  "Intdiff_11" ) 
#### Merge Data
diffscorereliable_g <- merge(LinkedMDAN_discord_AFI_g_9_m,LinkedMDAN_discord_AFI_g_10_m,by="ID",all=T)
diffscorereliable_g <- merge(diffscorereliable_g,LinkedMDAN_discord_AFI_g_11_m,by="ID",all=T)
matrix_g<-data.frame(diffscorereliable_g$Intdiff_9,diffscorereliable_g$Intdiff_10,diffscorereliable_g$Intdiff_11)
names(matrix_g)<-c("Age 9.5","Age 10.5","Age 11.5")

### First Born Boys
#### Import Data
LinkedMDAN_discord_AFI_b_9_m<-LinkedMDAN_discord_AFI_b_9[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_b_9_m) <-c("ID" ,"AFIMOMdiff_9", "INTMOMdiff_9","AFMKIDdiff_9","AFIKIDdiff_9","CSMOMdiff_9",  "Intdiff_9" ) 
LinkedMDAN_discord_AFI_b_10_m<-LinkedMDAN_discord_AFI_b_10[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_b_10_m) <-c("ID" ,"AFIMOMdiff_10", "INTMOMdiff_10","AFMKIDdiff_10","AFIKIDdiff_10","CSMOMdiff_10",  "Intdiff_10" ) 
LinkedMDAN_discord_AFI_b_11_m<-LinkedMDAN_discord_AFI_b_11[,c(2,57,67,69,71,73,92)]
names(LinkedMDAN_discord_AFI_b_11_m) <-c("ID" ,"AFIMOMdiff_11", "INTMOMdiff_11","AFMKIDdiff_11","AFIKIDdiff_11","CSMOMdiff_11",  "Intdiff_11" ) 
#### Merge
diffscorereliable_b <- merge(LinkedMDAN_discord_AFI_b_9_m,LinkedMDAN_discord_AFI_b_10_m,by="ID",all=T)
diffscorereliable_b <- merge(diffscorereliable_b,LinkedMDAN_discord_AFI_b_11_m,by="ID",all=T)
matrix_b<-data.frame(diffscorereliable_b$Intdiff_9,diffscorereliable_b$Intdiff_10,diffscorereliable_b$Intdiff_11)
names(matrix_b)<-c("Age 9.5","Age 10.5","Age 11.5")

### Export Cousin Cor
capture.output(xtable(cor_1star(matrix_b,2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_diffintreliable_b.tex"))
capture.output(xtable(cor_1star(matrix_g,2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_diffintreliable_g.tex"))
capture.output(xtable(cor_1star(matrix_z,2,.01,include.n=TRUE)),file=paste0(directory,"/Common/content/tables/table_diffintreliable_z.tex"))
@
<<measurement,include=FALSE,cache=TRUE>>=
# G1
## mplus
### Run
#### runModels(directory =paste0(directory,"/Common/content/mplusautomation/MplusG1"))
ModelSummaries_Mplus<-extractModelSummaries(target = paste0(directory,"/Common/content/mplusautomation/MplusG1"))
### Clean Up
ModelSummaries_Mplus[,4]<-NULL;ModelSummaries_Mplus[,3]<-NULL;ModelSummaries_Mplus[,2]<-NULL;ModelSummaries_Mplus[,1]<-NULL;ModelSummaries_Mplus[,23]<-NULL
### Export
capture.output(stargazer(ModelSummaries_Mplus,summary=FALSE,flip = TRUE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_gen1measurement.tex"))

# G2
## Mplus
### Run
#### runModels(directory =paste0(directory,"/Common/content/mplusautomation/MplusG2"))
g2loading_9 <- read.csv(paste0(directory,"/Common/content/data/g2loading_9.csv"), stringsAsFactors=FALSE)
g2loading_10 <- read.csv(paste0(directory,"/Common/content/data/g2loading_10.csv"), stringsAsFactors=FALSE)
g2loading_11 <- read.csv(paste0(directory,"/Common/content/data/g2loading_11.csv"), stringsAsFactors=FALSE)
### Extract
ModelSummaries_Mplusg2_9<-extractModelSummaries(target = paste0(directory,"/Common/content/mplusautomation/MplusG2/g_around_910811.out"))
ModelSummaries_Mplusg2_10<-extractModelSummaries(target = paste0(directory,"/Common/content/mplusautomation/MplusG2/g_around_1011912.out"))
ModelSummaries_Mplusg2_11<-extractModelSummaries(target = paste0(directory,"/Common/content/mplusautomation/MplusG2/g_around_11121013.out"))
### Clean
ModelSummaries_Mplusg2_9[,4]<-NULL;ModelSummaries_Mplusg2_9[,3]<-NULL;ModelSummaries_Mplusg2_9[,2]<-NULL;ModelSummaries_Mplusg2_9[,1]<-NULL;ModelSummaries_Mplusg2_9[,26]<-NULL;ModelSummaries_Mplusg2_9[,7]<-NULL;ModelSummaries_Mplusg2_9[,7]<-NULL;ModelSummaries_Mplusg2_9[,7]<-NULL

ModelSummaries_Mplusg2_10[,4]<-NULL;ModelSummaries_Mplusg2_10[,3]<-NULL;ModelSummaries_Mplusg2_10[,2]<-NULL;ModelSummaries_Mplusg2_10[,1]<-NULL;ModelSummaries_Mplusg2_10[,26]<-NULL;ModelSummaries_Mplusg2_10[,7]<-NULL;ModelSummaries_Mplusg2_10[,7]<-NULL;ModelSummaries_Mplusg2_10[,7]<-NULL

ModelSummaries_Mplusg2_11[,4]<-NULL;ModelSummaries_Mplusg2_11[,3]<-NULL;ModelSummaries_Mplusg2_11[,2]<-NULL;ModelSummaries_Mplusg2_11[,1]<-NULL;ModelSummaries_Mplusg2_11[,26]<-NULL;ModelSummaries_Mplusg2_11[,7]<-NULL;ModelSummaries_Mplusg2_11[,7]<-NULL;ModelSummaries_Mplusg2_11[,7]<-NULL
### Export
#### Age 9
capture.output(stargazer(ModelSummaries_Mplusg2_9,summary=FALSE,flip = TRUE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2_9measurement.tex"))
capture.output(stargazer(g2loading_9,summary=FALSE,flip = FALSE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2loading_9.tex"))

#### Age 10
capture.output(stargazer(ModelSummaries_Mplusg2_10,summary=FALSE,flip = TRUE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2_10measurement.tex"))
capture.output(stargazer(g2loading_10,summary=FALSE,flip = FALSE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2loading_10.tex"))

#### Age 11
capture.output(stargazer(ModelSummaries_Mplusg2_11,summary=FALSE,flip = TRUE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2_11measurement.tex"))
capture.output(stargazer(g2loading_11,summary=FALSE,flip = FALSE, dep.var.labels.include = FALSE),file=paste0(directory,"/Common/content/tables/table_g2loading_11.tex"))
@
<<results_regression,echo=FALSE, comment=NA, include=FALSE,cache=TRUE>>=
# Stargazer Plot defaults
font.size="small"
no.space=TRUE
style="demography"
column.sep.width = "1pt"
header = FALSE
column.labels = c("Mixed", "Daughters", "Sons")
model.numbers=FALSE
float=FALSE
ci =TRUE
single.row = TRUE
# 9.5
capture.output(stargazer(model.meanmoms_9,model.meanmoms.g_9,model.meanmoms.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_9.tex"))

capture.output(stargazer(model.meankids_9,model.meankids.g_9,model.meankids.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_9.tex"))

capture.output(stargazer(model.meanmomskids_9,model.meanmomskids.g_9,model.meanmomskids.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel","Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_9.tex"))

capture.output(stargazer(model.diffmoms_9,model.diffmoms.g_9,model.diffmoms.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen1 Mean Intel", "Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_9.tex"))

capture.output(stargazer(model.diffkids_9,model.diffkids.g_9,model.diffkids.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel", "Gen2 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_9.tex"))

capture.output(stargazer(model.diffmomskids_9,model.diffmomskids.g_9,model.diffmomskids.b_9,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel","Gen1 Mean Intel","Gen2 Dif Intel","Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_9.tex"))

# 10.5
capture.output(stargazer(model.meanmoms_10,model.meanmoms.g_10,model.meanmoms.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_10.tex"))

capture.output(stargazer(model.meankids_10,model.meankids.g_10,model.meankids.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_10.tex"))

capture.output(stargazer(model.meanmomskids_10,model.meanmomskids.g_10,model.meanmomskids.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel","Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_10.tex"))

capture.output(stargazer(model.diffmoms_10,model.diffmoms.g_10,model.diffmoms.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen1 Mean Intel", "Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_10.tex"))

capture.output(stargazer(model.diffkids_10,model.diffkids.g_10,model.diffkids.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel", "Gen2 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_10.tex"))

capture.output(stargazer(model.diffmomskids_10,model.diffmomskids.g_10,model.diffmomskids.b_10,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel","Gen1 Mean Intel","Gen2 Dif Intel","Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_10.tex"))

# 11.5
capture.output(stargazer(model.meanmoms_11,model.meanmoms.g_11,model.meanmoms.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_11.tex"))

capture.output(stargazer(model.meankids_11,model.meankids.g_11,model.meankids.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_11.tex"))

capture.output(stargazer(model.meanmomskids_11,model.meanmomskids.g_11,model.meanmomskids.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen1 Mean Intel","Gen2 Mean Intel"), 
                         dep.var.labels=c("Gen2 Mean AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_11.tex"))

capture.output(stargazer(model.diffmoms_11,model.diffmoms.g_11,model.diffmoms.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen1 Mean Intel", "Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_11.tex"))

capture.output(stargazer(model.diffkids_11,model.diffkids.g_11,model.diffkids.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel", "Gen2 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_11.tex"))

capture.output(stargazer(model.diffmomskids_11,model.diffmomskids.g_11,model.diffmomskids.b_11,
                         font.size=font.size,
                         no.space=no.space,
                         column.sep.width = column.sep.width,
                         header = header,
                         covariate.labels=c("Gen2 Mean AFI","Gen2 Mean Intel","Gen1 Mean Intel","Gen2 Dif Intel","Gen1 Dif Intel"),
                         dep.var.labels=c("Gen2 Differences in AFI"),
                         column.labels = column.labels,
                         model.numbers=model.numbers,
                         float=float,
                         ci =ci,
                         single.row = single.row),
               file=paste0(directory,"/Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_11.tex"))

# Summary Stats by Race and Gender
capture.output(latex(
  tabular((Heading()*( 1 + RowFactor(CSEX)+ RowFactor(RACE) + RowFactor(CSEX)*RowFactor(RACE))~(Heading("AFI")*AFI_G2*(Mean + Sd))), 
          data = Gen2AFIConcor_2),
  numeric.dollar=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_AFIRACEGENDER.tex"))
@
<<setup_tables,include=FALSE,cache=TRUE>>=
# Summary Stats G1
## Manage data
### Import
df <- merge(Gen1AFIConcor,birthweight_fb,by="MCASEID",all=T)
nlsyextract <- read.csv(paste0(directory,"/Common/content/data/nlsyextract.csv"))
###
### Clean
nlsyextract$IDC1[nlsyextract$IDC1 <= 0  ] <- 0
nlsyextract$IDC1[nlsyextract$IDC1 > 0  ] <- 1
nlsyextract$SiblinginSample <-nlsyextract$IDC1
nlsyextract$IDC1 <- NULL
nlsyextract[nlsyextract< 0  ] <- NA
nlsyextract$month<-nlsyextract$Q1.3_A.M 
nlsyextract$month<- paste("0",nlsyextract$Q1.3_A.M,sep="")
nlsyextract$month[nlsyextract$month == 010 ] <- 10
nlsyextract$month[nlsyextract$month == 011 ] <- 11
nlsyextract$month[nlsyextract$month == 012 ] <- 12
nlsyextract$Age<-as.numeric(as.Date(paste(df$YRB,"-06-15",sep=""), format='%Y-%m-%d')-as.Date(paste("19",nlsyextract$Q1.3_A.Y,"-",nlsyextract$month,"-15",sep=""), format='%Y-%m-%d'))/365.25
summarystats_g1<-data.frame(nlsyextract$SiblinginSample,nlsyextract$Age,df$Intel_G1,df$AFI_G1,nlsyextract$SAMPLE_SEX)
###
### Sibling in sample?
#### Gen2AFIConcor$MCASEID<-birthweight$MCASEID
#### data=Gen2AFIConcor
#### data$dup<-0
#### for (x in 1:length(data[,1])) {
#### for (y in 1:length(data[,1])) {
#### if (data$MCASEID[x]==data$MCASEID[y]) data$dup[x]= data$dup[x]+1 }}
#### write.table(data, file = paste0(directory,"/Common/content/data/g2sibinsample.csv"), append = FALSE)
Gen2AFIConcor<- read.table(paste0(directory,"/Common/content/data/g2sibinsample.csv"), header=TRUE, quote="\"")
Gen2AFIConcor$SiblinginSample<-1
Gen2AFIConcor$SiblinginSample[Gen2AFIConcor$dup == 1 ]<- 0
summarystats_g2 <- merge(categories_g2,Gen2AFIConcor,by="CASEID",all=T)
summarystats_g2$zIntel_G2<-scale(summarystats_g2$Intel_G2)
summarystats_g2 <- merge(summarystats_g2,birthweight_2,by="CASEID",all=T)
##
## Capture output
capture.output(latex(
  tabular(((Heading("Mother Age at Birth")*nlsyextract.Age + Heading("Standardized AFQT")*df.Intel_G1 + Heading("AFI")*df.AFI_G1 ) ~ Heading("Has Sibling in NLSY?")*(Mean + Sd)*Heading()*( 1 + RowFactor(nlsyextract.SiblinginSample) )), data = summarystats_g1),
  numeric.dollar=TRUE,na.blank=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_sibinsampleg1.tex"))

capture.output(latex(
  tabular(Heading("Year Born")*YRB + Heading("Standardized Ability Score")*zIntel_G2 + Heading("AFI")*AFI_G2 ~ (Heading("Has Sibling in NLSY?")*(Mean + Sd)*Heading()*(1 + RowFactor(SiblinginSample))), data = summarystats_g2),
  numeric.dollar=TRUE,na.blank=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_sibinsampleg2.tex"))
@

%%%%%%%%%%%% Introduction %%%%%%%%%%%%
%\section{Introduction}
\section{ }\vspace{-.8cm}
% Introduce the Major Problem
\input{../Common/content/writing/introduction.txt}
% Current Study
\subsection{Current Study, Summary}
\input{../Common/content/writing/current_study.txt}

%%%%%%%%%%%% Method %%%%%%%%%%%%
\section{Method}
\subsection{Research Design}
\input{../Common/content/writing/design.txt}
%
% Sample
\subsection{Sample}
\input{../Common/content/writing/subjects.txt}
%
% Tetrads
\subsection{Tetrads}
\input{../Common/content/writing/tetrad.txt}
%
%%%%%%%%%%%% Measurement %%%%%%%%%%%%
% Age at First Intercourse
\subsection{Measures}
\input{../Common/content/writing/AFI.txt}
\noindent\begin{minipage}{\linewidth}
\captionof{figure}{Smoothed Density Plots of AFI by Sex and Race}
\label{plot_afi_by_race_sex}
\begin{center}
<<plot_afi_by_race_sex,fig=TRUE,echo=FALSE,cache=TRUE,echo=FALSE, fig.height=5*.6, fig.width=5,out.width=".8\\paperwidth",warning=FALSE>>=
  # Color Blind Palette 
  cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7","#000000")
  # Data Prep
  dataframe<-data.frame(Gen2AFIConcor_2,Gen2AFIConcor$Intel_G2)
  names(dataframe)<- c("CASEID","MCASEID","RACE","SEX","AFI","Intel")

  ## Race by Sex factor
  dataframe$RACEbySEX<-"0"
  ###
  ### Hispanic
  #### Male
  dataframe$RACEbySEX[dataframe$RACE=='Hispanic' & dataframe$SEX=='Male']<-"Hispanic Male"
  ####
  #### Female
  dataframe$RACEbySEX[dataframe$RACE=='Hispanic'&dataframe$SEX=='Female']<-"Hispanic Female"
  ###
  ### Black
  #### Male
  dataframe$RACEbySEX[dataframe$RACE=='Black'&dataframe$SEX=='Male']<-"Black Male"
  ####
  #### Female
  dataframe$RACEbySEX[dataframe$RACE=='Black'&dataframe$SEX=='Female']<-"Black Female"
  ###
  ### Non-Black, Non-Hispanic
  #### Male
  dataframe$RACEbySEX[dataframe$RACE=='Non-Black, Non-Hispanic'&dataframe$SEX=='Male']<-"White Male"
  ####
  #### Female
  dataframe$RACEbySEX[dataframe$RACE=='Non-Black, Non-Hispanic'&dataframe$SEX=='Female']<-"White Female"
  ###
  ### Unknown Race
  dataframe$RACEbySEX[dataframe$RACEbySEX=='0']<-"Unknown"
  dataframe$RACE<-revalue(dataframe$RACE, c("Non-Black, Non-Hispanic"="White"))
  #summary(as.factor(dataframe$RACEbySEX))

  ggplot(dataframe, aes(x=AFI, colour=RACE, group=RACEbySEX, linetype = SEX))+ geom_density(fill=NA,adjust=2)+ theme_bw() + theme(axis.text=element_text(size=10), axis.title=element_text(size=12), plot.title=element_text(size=12)) +  theme(legend.position="right",  legend.justification=c(1, 0.5)) +   xlab("AFI") + ylab("Density") +scale_colour_manual(values= c("#56B4E9", "#009E73", "#CC79A7")) + coord_cartesian(ylim=c(0,0.30), xlim=c(12,25))
  @
\end{center}
\end{minipage}
%
% Intelligence
\input{../Common/content/writing/intelligence.txt}
%
% Difference Score Reliability
\subsection{Reliability of Difference Scores}
<<calculationalphadif,cache=TRUE,include=FALSE>>=
# Intelligence
sigmax = sqrt(1.1881)
sigmay = sigmax
rxx = .87
ryy = rxx
rxy=.67
r.diff.intelligence <- (sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) /(sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay)
#AFI
sigmax = sqrt(4.41)
sigmay = sigmax
rxx = .76
ryy = rxx
rxy=.099
r.diff.afi <- (sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) / (sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay)
@
Our design assumes that the difference scores of our measures are reliable. We've reported the test-retest reliability of Gen2 intelligence and Gen1 AFI in earlier sections. Here, we report the test-retest reliability of the pairwise differences of those measures. 
\subsubsection{AFI} Comparing sibling differences in AFI as reported in 1983 and 1984 (n = 783 pairs) we found a strong correlation (r = .76). The sample of sibling pairs with complete information in 1985 was too small (n = 12 pairs) to compare to the other two years. Regardless, sibling differences in self-reported AFI appear reliable. Because we could not estimate test-retest reliabilities for Generation 2, we calculated the reliability using Lord's equation (\citeyear{Lord1963}). Generation 2 Mean AFI difference scores were also reliable (r = \Sexpr{r.diff.afi}) and comparable to Generation 1 sibling differences.

\subsubsection{Intelligence} Cousin differences in intelligence as assessed at ages 9.5, 10.5, and 11.5 were correlated using three different linking methods (Mixed, Daughers, Sons). Reliabilities across linking methods was consistent and high (min r = .86; max r = .95). However, we could not estimate test-retest reliabilities for Generation 1; we calculated the reliability using Lord's equation (\citeyear{Lord1963}). The calculated reliability of Generation 1's differences in AFQT was \Sexpr{r.diff.intelligence}, acceptable, but lower than the empirical correlation we derived for cousin differences.

%%%%%%%%%%%% Results %%%%%%%%%%%%
\section{Results}
\input{../Common/content/results/results.txt}
%
% Between Family
\subsection{Between-Family Analyses}
\input{../Common/content/results/results_between.txt}
%
% Within Family
\subsection{Within-Family Analyses}
\input{../Common/content/results/results_within.txt}
%
%%%%%%%%%%%% Discussion %%%%%%%%%%%%
\section{Discussion}
\input{../Common/content/discussion/discussion.txt}

%%%%%%%%%%%% References %%%%%%%%%%%%
\bibliography{Projects-AFI}
  %%%%%%%%%%%% Appendix %%%%%%%%%%%%
   \begin{landscape} \appendix\label{appen}
  %
  % Age 10.5 Replication
  \section{Age 10.5 Replication}\label{appen10}
  %% Between Family

  \subsection{Between Family Analyses}
  %%%
  %%% Mom -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Mean Gen1 Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{22}{../Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen2 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{22}{../Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Mom Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Mean Joint Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{23}{../Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%
    %% Within Family
  \subsection{Within Family Analyses}
  %%%
  %%% Mom -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen1 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{24}{../Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen2 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{24}{../Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Mom Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Joint Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{26}{../Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_10.tex}
  \end{longtable}
  % Age 11.5 Replication
  \section{Age 11.5 Replication}\label{appen11}
  %% Between Family
  \subsection{Between Family Analyses}
  %%% Mom -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Mean Gen1 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{22}{../Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen2 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{22}{../Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Mom Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Mean Joint Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{23}{../Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%
    %% Within Family
  \subsection{Within Family Analyses}
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen1 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{24}{../Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen2 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{24}{../Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Mom Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Dif Joint Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{26}{../Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_11.tex}
  \end{longtable}
  \end{landscape}
  %
\end{document}