### 2015-12-06 Manuscript/Manuscript.tex

\documentclass[a4paper,man,apacite,natbib,12pt,longtable]{apa6}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{../Common/style/Mason}
\usepackage{listings}
\usepackage{inconsolata}

%%%%%%%%%%%% Title %%%%%%%%%%%%
\title{Is the Link Between Intelligence and Age at First Intercourse Causal or Correlational? \\ A Cross-Generational Sibling Comparison Design Using the NLSY}
\shorttitle{Intelligence and AFI}
%
% Authors and Affiliations
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}
%
% Author Note
\authornote{\input{../Common/content/title/authornote.txt}}
%
% Abstract
\abstract{Last compile was \today\ at \currenttime\\
\input{../Common/content/title/abstract_empreport.txt}}
%
% Key Words
\keywords{Age at first intercourse; Cross-sectional Data; Discordant Sibling Design; Intelligence; Quasi-Experimental; Siblings}
%
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\maketitle
%
%%%%%%%%%%%% R %%%%%%%%%%%%

%%%%%%%%%%%% Introduction %%%%%%%%%%%%
%\section{Introduction}
\section{ }\vspace{-.8cm}
% Introduce the Major Problem
\input{../Common/content/writing/introduction.txt}
% Current Study
\subsection{Current Study, Summary}
\input{../Common/content/writing/current_study.txt}

%%%%%%%%%%%% Method %%%%%%%%%%%%
\section{Method}
\subsection{Research Design}
\input{../Common/content/writing/design.txt}
%
% Sample
\subsection{Sample}
\input{../Common/content/writing/subjects.txt}
%
% Tetrads
\subsection{Tetrads}
\input{../Common/content/writing/tetrad.txt}
%
%%%%%%%%%%%% Measurement %%%%%%%%%%%%
% Age at First Intercourse
\subsection{Measures}
\input{../Common/content/writing/AFI.txt}
\noindent\begin{minipage}{\linewidth}
\captionof{figure}{Smoothed Density Plots of AFI by Sex and Race}
\label{plot_afi_by_race_sex}
\begin{center}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=.8\paperwidth]{figure/plot_afi_by_race_sex-1}

\end{knitrout}
\end{center}
\end{minipage}
%
% Intelligence
\input{../Common/content/writing/intelligence.txt}
%
% Difference Score Reliability
\subsection{Reliability of Difference Scores}

Our design assumes that the difference scores of our measures are reliable. We've reported the test-retest reliability of Gen2 intelligence and Gen1 AFI in earlier sections. Here, we report the test-retest reliability of the pairwise differences of those measures. 
\subsubsection{AFI} Comparing sibling differences in AFI as reported in 1983 and 1984 (n = 783 pairs) we found a strong correlation (r = .76). The sample of sibling pairs with complete information in 1985 was too small (n = 12 pairs) to compare to the other two years. Regardless, sibling differences in self-reported AFI appear reliable. Because we could not estimate test-retest reliabilities for Generation 2, we calculated the reliability using Lord's equation (\citeyear{Lord1963}). Generation 2 Mean AFI difference scores were also reliable (r = 0.734) and comparable to Generation 1 sibling differences.

\subsubsection{Intelligence} Cousin differences in intelligence as assessed at ages 9.5, 10.5, and 11.5 were correlated using three different linking methods (Mixed, Daughers, Sons). Reliabilities across linking methods was consistent and high (min r = .86; max r = .95). However, we could not estimate test-retest reliabilities for Generation 1; we calculated the reliability using Lord's equation (\citeyear{Lord1963}). The calculated reliability of Generation 1's differences in AFQT was 0.606, acceptable, but lower than the empirical correlation we derived for cousin differences.

%%%%%%%%%%%% Results %%%%%%%%%%%%
\section{Results}
\input{../Common/content/results/results.txt}
%
% Between Family
\subsection{Between-Family Analyses}
\input{../Common/content/results/results_between.txt}
%
% Within Family
\subsection{Within-Family Analyses}
\input{../Common/content/results/results_within.txt}
%
%%%%%%%%%%%% Discussion %%%%%%%%%%%%
\section{Discussion}
\input{../Common/content/discussion/discussion.txt}

%%%%%%%%%%%% References %%%%%%%%%%%%
\bibliography{Projects-AFI}
  %%%%%%%%%%%% Appendix %%%%%%%%%%%%
   \begin{landscape} \appendix\label{appen}
  %
  % Age 10.5 Replication
  \section{Age 10.5 Replication}\label{appen10}
  %% Between Family

  \subsection{Between Family Analyses}
  %%%
  %%% Mom -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Mean Gen1 Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{22}{../Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen2 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{22}{../Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Mom Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Mean Joint Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{23}{../Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%
    %% Within Family
  \subsection{Within Family Analyses}
  %%%
  %%% Mom -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen1 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{24}{../Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen2 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{24}{../Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_10.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Mom Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Joint Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_10}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Difference in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{26}{../Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_10.tex}
  \end{longtable}
  % Age 11.5 Replication
  \section{Age 11.5 Replication}\label{appen11}
  %% Between Family
  \subsection{Between Family Analyses}
  %%% Mom -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Mean Gen1 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Mom_Intelligence_Mean_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{22}{../Common/content/tables/table_Mean_Mom_Intelligence_Mean_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen2 Mean Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Child_Intelligence_Mean_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{22}{../Common/content/tables/table_Mean_Child_Intelligence_Mean_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Mom Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Mean Joint Intelligence $\rightarrow$ Gen2 Mean AFI}\label{table_Mean_Joint_Intelligence_Mean_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Average of Gen2 AFI} \\ 
  \cline{2-4}
  \partialinput{10}{23}{../Common/content/tables/table_Mean_Joint_Intelligence_Mean_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%
    %% Within Family
  \subsection{Within Family Analyses}
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen1 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Mom_Intelligence_Dif_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{24}{../Common/content/tables/table_Dif_Mom_Intelligence_Dif_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Gen2 Dif Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Child_Intelligence_Dif_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{24}{../Common/content/tables/table_Dif_Child_Intelligence_Dif_Child_AFI_11.tex}
  \end{longtable}\pagebreak
  %%%
  %%% Mom Child -> Child
  \begin{longtable}{@{\extracolsep{5pt}}lccc} 
  \caption{Dif Joint Intelligence $\rightarrow$ Gen2 Dif AFI}\label{table_Dif_Joint_Intelligence_Dif_Child_AFI_11}
  \\[-1.8ex]\hline 
  \hline \\[-1.8ex] 
  & \multicolumn{3}{c}{\textit{Dependent variable:} Differences in Gen2 Mean AFI} \\ 
  \cline{2-4}
  \partialinput{10}{26}{../Common/content/tables/table_Dif_Joint_Intelligence_Dif_Child_AFI_11.tex}
  \end{longtable}
  \end{landscape}
  %
\end{document}