### 2015-12-12 Common/content/title/authornote.tex
tables

{\small S. Mason Garrison, Department of Psychology and Human Development, Vanderbilt University; Joseph Lee Rodgers, Department of Psychology and Human Development, Vanderbilt University.

This manuscript is based on longitudinal data collected as part of the National Longitudinal Survey of Youth 1979, (NLSY79). These data are publicly available; a bibliography of articles using these surveys can be found at \url{https://nlsinfo.org/bibliography-start} This material is based upon work that has been supported by the National Institute of Health under Grant No. (R01-HD065865) and the National Science Foundation Graduate Research Fellowship Program under Grant No. (DGE-1445197), and various means of institutional support from the following universities: University of British Columbia, University of Oklahoma \& Vanderbilt University. The aforementioned funding sources did not have any input in the production of this article.

Correspondence concerning this article should be addressed to S. Mason Garrison, Department of Psychology and Human Development, Vanderbilt University, Nashville, TN. Contact: s.mason.garrison@gmail.com}