### 2015-12-12 Common/content/writing/results.tex

We examined the relationship between AFI and intelligence using two designs: a between-family design, and a within-family design (which also includes between-family variance in the differences between the family means) . The results are organized bythose two designs. The between-family analyses report the relationships between the within-family average AFI and various measures of ability. The within-family analyses replicate the between-family analyses using within-family difference scores by testing whether differences in AFI can be explained by differences in various measures of ability.