### 2016-06-17 Common/content/writing/results.tex

We examined the relationship between AFI and intelligence using two designs: a between-family design, and a combination between- and within-family design (which includes between-family variance in the differences between the family means, and within-family variance in the sibling/cousin differences). The results are organized by those two designs. The between-family analyses report the relationships between the within-family average AFI and various measures of ability. The combination between-family/within-family analyses add to the between-family analyses the difference scores, testing whether differences in AFI can be explained by differences in various measures of ability, controlling for the between-family variance.